### 2019-12-27 manuscript/main.Rmd

---
title: Historical aerial surveys map long-term changes of forest cover and structure in the central Congo Basin

author: 
  - name: Koen Hufkens
    affil: 1,2,*
    orcid: 0000-0002-5070-8109
  - name: Thalès de Haulleville
    affil: 3
  - name: Elizabeth Kearsley
    affil: 1
  - name: Kim Jacobsen
    affil: 2,3
  - name: Hans Beeckman
    affil: 3
  - name: Piet Stoffelen
    affil: 4
  - name: Filip Vandelook
    affil: 4
  - name: Sofie Meeus
    affil: 4
  - name: Michael Amara
    affil: 5
  - name: Leen Van Hirtum
    affil: 5
  - name: Jan Van den Bulcke
    affil: 1
  - name: Hans Verbeeck
    affil: 1
  - name: Lisa Wingate
    affil: 2

affiliation:
  - num: 1
    address: Ghent University, Ghent, Belgium
  - num: 2
    address: INRA, UMR ISPA, Villenave d’Ornon, France
  - num: 3
    address: Royal Museum for Central Africa, Tervuren, Belgium
  - num: 4
    address: Botanic Garden Meise, Meise, Belgium
  - num: 5
    address: National Archives of Belgium, Brussels, Belgium

correspondence: koen.hufkens@gmail.com
journal: remote sensing
type: article
status: submit
bibliography: bibliography.bib

header-includes: >
  \usepackage{wrapfig}
  \usepackage{graphics}
  \usepackage{setspace}
  \doublespacing
  \usepackage{gensymb}
  \usepackage{booktabs}

abstract: |
   Land Use and Land Cover change (LULCC) of African rainforest contribute to global carbon emissions. Yet, most historical estimates of LULCC and their carbon emissions rely on non-spatially explicity data in the pre-satellite era (< 1972). Past studies have expanded this pre-satellite time frame with declassified satellite surveillance data or aerial surveys, but none cover the Congo Basin. Here, we use historical aerial survey photos to map the extent and structure of LULCC around Yangambi, central Congo Basin, in 1958. Our study leveraged Structure-from-Motion to generate a large orthomosaic covering 828 km^2^, counting 733 million pixels, at a sub meter resolution and geo-referenced to ~4.7 $\pm$ 4.3 m. Primary forest in the orthomosaic was classified with a Convolutional Neural Network Deep Learning approach. We used these data to quantify LULCC, landscape and canopy complexity and scale above ground biomass between historical and contemporary condition. We show a shift from historical highly structured industrial deforestation in 1958 (162 km^2^), to contemporary smallholder farming clearing (88 km^2^), increasing landscape fragmentation and forest edges. Efforts to quantify canopy texture and their link to carbon storage had limited to no success. Our analysis provides an insight in key spatial and temporal patterns of deforestation and reforestation at a multi-decadal scale and provide a historical context to land-cover and land-use change spatial patterns for past and ongoing field research in the area.
keywords: |
  Aerial survey, data recovery, CNN, Deep Learning, SfM, Congo Basin
acknowledgement: |
  This research was supported through the Belgian Science Policy office COBECORE project (BELSPO; grant BR/175/A3/COBECORE). KH acknowledges funding from the European Union Marie Skłodowska-Curie Action (project number 797668). I'm grateful for the support of Thales D'Haulleville volunteering his time in scanning the images.
authorcontributions: |
  K.H. conceived and designed the study, analyzed the data, prepared figures and tables, authored, reviewed and approved the final draft of the manuscript. T.d.H. scanned all image data. E.K. and T.d.H. provided plot based AGC estimates. T.D., K.J., E.K, H.B., P.S., F.V.,S.M., M.A., J.V.D.B., H.V and L.W. reviewed the final manuscript.
conflictsofinterest: |
  The authors declare no conflict of interest. The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, an in the decision to publish the results.
output: rticles::mdpi_article
---

```{r setup, include=FALSE, message=FALSE}
  knitr::opts_chunk$set(echo = TRUE)
  library(orthodrc)
  library(raster)
  library(ggplot2)
  library(ggthemes)
  library(tidyverse)
  library(knitr)
  library(kableExtra)
  library(landscapemetrics)

  # set kable NA values to blank
  options(knitr.kable.NA = '-')

  # landscape stats function
  landscape_stats <- function(r){

  df <- lsm_p_perim(r) %>%
    filter(class == 1) %>%
    select(-c(metric)) %>%
    rename(perimeter = value)

  df$perimeter <- df$perimeter

  area <- lsm_p_area(r) %>%
    filter(class == 1) %>%
    select(value) %>%
    rename(area = value)

  frac <- lsm_p_frac(r)  %>%
    filter(class == 1) %>%
    select(value) %>%
    rename(fractal_index = value)

  df <- data.frame(df, area = area, frac = frac)
  df$pa_ratio <- df$perimeter/(df$area*(100*100))

  df <- df %>%
    filter(area > 1) %>%
    summarize(
      perimeter = paste0(
        round(mean(perimeter)),
        " $\\pm$ ",
        round(sd(perimeter))),
      area = paste0(
        round(mean(area)),
        " $\\pm$ ",
        round(sd(area))),
      ratio = paste0(
        round(mean(pa_ratio),3),
        " $\\pm$ ",
        round(sd(pa_ratio),3)),
      fractal = paste0(
        round(mean(fractal_index),3),
        " $\\pm$ ",
        round(sd(fractal_index),3))
    )

  return(df)
}

  # read data
  change_map <- raster("../data/orthomosaic/yangambi_forest_cover_difference_1958_2000.tif")
  # 
  # read site data
  psp <- read.table("../data/surveys/site_characteristics.csv",
                    header = TRUE,
                    sep = ",")

  # patch metrics
  # select classes into their respective time frame
  historical <- (change_map == 2 | change_map == 4)
  contemporary <- (change_map == 3 | change_map == 4)

  # reproject to meter based mercator projection
  # should be fine around the equator
  historical <- projectRaster(historical,
                              crs = "+init=epsg:3395",
                              method = "ngb")

  contemporary <- projectRaster(contemporary,
                              crs = "+init=epsg:3395",
                              method = "ngb")

  patch_metrics <- rbind(landscape_stats(historical),
                 landscape_stats(contemporary))

  row.names(patch_metrics) <- c("historical","contemporary")

  # read change stats
  change_stats <- read.csv2("../data/surveys/lulcc_change_stats.csv",
                            header = TRUE,
                            sep = ",")
  change_stats[1:2,3] <- NA

  row.names(change_stats) <- c(
   "Forest",
   "  -  of which edges",
   "Regrowth > 1958 loss",
   "Loss > 2000",
   "Loss > 1958 (permanent)")

```

# Introduction

The functioning of a tropical forest, and its capacity to sequester carbon, is heavily influenced by anthropogenic disturbances such as selective logging, clear felling, or slash and burn practices, all increasing forest fragmentation. This increased pace in the loss of rainforest cover [@butsic2015] is predicted to negatively impact the function of these tropical forests, and their future capacity to sequester carbon, as anthropogenic disturbances such as selective logging, clear felling, or slash and burn practices collectively contribute to increasing forest fragmentation.

These forest fragments and their edges contain fewer large trees [@laurance2000], in part due to their hotter and brighter microclimates [@magnago2015]. Similarly, species composition and biodiversity [@poorter2006] are negatively affected. Spatial disturbances also have an explicit temporal component, where the time since disturbance affects the strength of the effects [@didham1999]. Deforestation and gap formation also influence the overall canopy structure, where pioneer species grow taller and with narrower crowns [@poorter2006; @peres2006] affecting the forest light regime and carbon dynamics [@stark2012]. As such, anthropogenic factors through land-use and land cover change (LULCC) heavily affect forest structure and functioning [@fauset2017; @brinck2017; @didham1999]. In the tropics the majority of emissions originate from deforestation and forest degradation [@hansen2013; @houghton2012; @tyukavina2015] contributing to 10-15% of the total global emissions [@vanderwerf2009] and concomitant losses in biodiversity losses [@barlow2016; @vandeperre2018] and how LULCC is evolving over time. Although the African rainforest is the second largest on Earth and covers ~630 million ha, representing up to 66 Pg of carbon storage [@Lewis2009b], and currently loses forest at an increasing pace [@butsic2015]. Yet, the tropics is now becoming a large source of CO_2 to the atmosphere as deforestation and forest degradation [@hansen2013; @houghton2012; @tyukavina2015] contribute up to 10-15% of the total global CO2 emissions [@vanderwerf2009]. Furthermore, recent estimates show that 31% of recent carbon emissions are caused by edge effects alone [@brinck2017]. In this context, there are very few long term forest inventory plots that presently exist in the central Congo Basin [@Verbeeck2011a; @lewis2013]. This further limits the accuracy of current biomass estimates and challenges our understanding of how forest structure and function changes in the short and long term in these ecosystems [@Kearsley2013; @Doetterl2015].

Historical estimates of LULCC, and their carbon emissions, exist [@houghton2012; @ramankutty1999; @defries2002; @sader1988], however they generally rely on non-spatially explicit data in the pre-satellite era (< 1972) [@houghton2010]. These analysis often use FAO statistics, historical land-use sources, and/or population and socio-economic data [@houghton2012], as limited earlier spatial data sources are available [@hansen2013; @achard2014]. For example, the @houghton2010 long-term study used FAO and national censuses alongside a bookkeeping model to estimate carbon emissions from LULCC. Other studies use remote sensing data, but are limited in time to recent decades [@defries2002; @houghton2012]. These proxy based studies have a limited view on long-term direct and indirect spatial effects of deforestation. As such, both the spatial structure and the fragmentation pattern of a forest combined with its life history has a profound effect on carbon emssisions. However, in this context remote sensing assessments provide more accurate spatial information to help estimate carbon emissions [@mitchard2018].

Past studies expanded knowledge during the pre-satellite time frame through the use of now declassified satellite surveillance data to map deforestation in the US, central Brazil, and post world war II Germany [@song2015; @nita2018]. In some cases these records could extend back in time even further using historical aerial survey images. In northern Europe especially these data are well known and used, for example to map the extent of forest cover across ~100 km^2^ regularly, and at times using automated workflows [@giordano2018]. Survey data across the African continent is less common or inaccessible or both. Some studies do exist, both @buitenwerf2012 and @okeke2006 used aerial survey images to map vegetation changes in South African savannas and the Adulam Nature Reserve, respectively. Whilst @frankl2015 used 20 aerial survey images to map the Ethiopian highlands of 1935. High resolution aerial or surveillance satellite images provide scientists tools to monitor forest extent and structure over long time frames. Moreover, canopy texture from aerial survey images has been successfully linked to above ground biomass [@Ploton2012; @Couteron2005; @Barbier2010]. Thus using texture based metrics obtained from (high resolution) remote sensing data opens up possibilities to link historical aerial footage to carbon density across large spatial extents [@Couteron2005].

Given the impact of LULCC on the structure and functioning of tropical forests, and their influence on both carbon dynamics [@Kearsley2013] and biodiversity [@vandeperre2018], accurate long term reporting of historical forest cover in the pre-satellite era is an imperative [@willcock2016]. In this study we use a combination of historical aerial photography and contemporary remote sensing data to map long-term changes in the extent and structure of the tropical forest surrounding Yangambi, in the central Congo Basin at the beginning of 1958 and the start of the antrophocene [@lewis2015]. Our analysis provides a historical insight into important LULCC spatial patterns in Yangambi, such as fragmentation and edge complexity. Our analysis further contextualises the influence of changes in the forest's life history on past and current research into Above Ground Carbon (AGC) storage [@Kearsley2013] and biodiversity [@vandeperre2018] in the central Congo Basin.

# Methods

## Historical data acquisition

Data for the central Congo Basin region, surrounding Kisangani, were collected in several flights during the dry season of 1958 and 1959 (from 8/01/1958 to 20/02/1958 and from 28/12/1958 to 9/01/1959 respectively, see Appendix Figure 1) to generate topographic maps of the area, supervised by the "Institut Géographique du Congo Belge" in Kinshasa (then Léopoldville). Black-and-White infrared images (0.4 - 0.9 $\mu$m) were gathered along flight paths running mostly from west to east, between 9 - 11h local time. Along a flight path continuous images were taken using a Wild Heerburgg RC5a (currently Leica Geosystems) with an Aviogon lens assembly (114.83mm / f 5.6, with a 90\degree view angle) resulting in square photo negative of 180 by 180 mm. Flights were flown at an average absolute altitude of ~5200 m above sea level, covering roughly 18 530 km^2^ at an approximate scale of 1/40 000. The use of the integrated autograph system ensured timely acquisition of pictures with a precise overlap (~1/3) between images. This large overlap between images together with flight parameters would allow post-processing, using stereographs, to create accurate topographic maps. Original data from this campaign are stored in the [Royal Museum for Central Africa](https://www.africamuseum.be/en) in Tervuren, Belgium (Figure 1).

```{r echo=FALSE, out.width='100%', fig.cap="A box of historical aerial photographs (left) and a single aerial photograph (right) showing part of the Congo river. Note the meta-data provided in the margins of the image such as acquisition time and fleight height and the solar glare on the Congo river."}
knitr::include_graphics('./figures/print_examples.jpg')
```

## Site selection

```{r echo=FALSE, out.width='100%', message=FALSE, warning=FALSE, fig.cap="Overview of the historical flight paths during aerial photo acquisition and ancillary data used in this study. The bounding box of the orthomosaic data presented in this study is shown as a rectangle (23x36 km). The outline of a recent high-resolution Geo-eye panchromatic image is shown as a dashed dark grey rectangle (10x10km). The location of various permanent sampling plots are shown as x, +, and open squares and triangles for the mixed, mono-dominant and edge plots respectively. The grey polygon delineates the current day Yangambi Man and Biosphere reserve. The inset, top right, situates the greater Yangabmi region with the DR Congo. The full flight plan and details are shown in Appendix Figure 1 and 2"}
knitr::include_graphics('./figures/flight_paths.png')
```

We prioritised flight paths and images that contained current day permanent sampling plots, larger protected areas, and past agricultural and forest research facilities (Figure 2). This selection provides a comprehensive mapping of the Yangambi area and the life history of the forest surrounding it. Thereafter, we selected flight paths 1 through 11 for digitization. From this larger dataset of 334 images we selected 74 survey images for orthomosaic compositing and further analysis. All the selected images stem from the flight campaign made during January and February of 1958. The area includes the Yangambi village, 20 contemporary permanent sampling plots [@Kearsley2013], past and present agricultural experimental plots [@bauters2015] and large sections of the Yangambi [UNESCO Man and Biosphere reserve](http://www.unesco.org/new/en/natural-sciences/environment/ecological-sciences/biosphere-reserves/africa/democratic-republic-of-the-congo/yangambi/) surrounding to the west and east of the village. Although not formally mosaicked we provided a full dataset of pre-processed images using the cropping and normalization routines described below. The latter data was not used in subsequent LULCC analysis, but has been archived and made available to the public separately (see code & data availability statement below).

## Digitization and data processing

All selected images, covering the Yangambi area, were contact prints as original negatives of the prints were not available. Images were scanned at a resolution exceeding their original resolution (or grain) at the maximal physical resolution of an Epson A3 flatbed scanner (i.e. 2400 dpi or 160MP per image) and saved as lossless tiff images. Data were normalized using contrast limited histogram equalization [@zuiderveld1994] with a window size of 32 and a clip limit of 1.5. Fiduciary marks were used to rectify and downsample the images into square 7700x7700 pixel images (~1200 dpi, 81MP). This resulted in a dataset with digital images at a resolution that remained above the visible grain of the photographs, whilst the reduced image size facilitated easier file handling and processing speed.

Data was processed into a georeferenced orthomosaic using a Structure from motion (SfM, @ullman1979) approach implemented in [Agisoft Metashape](http://www.agisoft.com) version 1.5.2 (Agisoft LLC, St. Petersburg, Russia). An orthomosaic corrects remote sensing data to represent a perfectly downward looking image, free from perspective distortions due to topography and camera tilt. Using the SfM technique features, areas in images with a large degree of similarity, are matched across various images to reconstruct a three dimensional scene (topography) from two-dimensional image sequences. During the SfM analysis we masked clouds, glare or large water bodies such as the Congo river.

We calculated the orthomosaic using a low resolution point cloud and digital elevation map (DEM). Additional ground control points were provided to assist in the referencing of image and constrain the optimization routine used in the SfM algorithm. Ground control points consisted of rooftop edges of permanent structures which could be verified in both old and new aerial imagery (i.e. ESRI World Imagery). Although clouds were removed during the SfM routine we did not mask all clouds in the final orthomosaic to maximize forest coverage. The final scene was cropped, to provide consistent wall to wall coverage of the reconstructed scene. The orthomosaic was exported as a geotiff for further georeferencing in QGIS [@qgisdevelopmentteam2018] using the georeferencer plugin (version 3.1.9) and additional ESRI World Imagery high resolution reference data. We used 3^th^ degree polynomial and 16 ground control points to correct the final image. Ground control points, raw image data and final processed image are provided in addition to measures of uncertainty such as mean root mean squared (RMSE), mean and median error across all ground control points. All subsequent analysis are executed on the final geo-referenced orthomosaic or subsets of it.

## Land-Use and Land-Cover Change

### Classifying forest cover

We automatically delineated all natural forest in the historical data, thus excluding tree plantations, thinned or deteriorated forest stands, fields and buildings. We used the Unet Convolutional Neural Net (CNN, @ronneberger2015) architecture implemented in Keras [@chollet2015] with an efficientnetb3 backbone [@yakubovskiy2019] running on TensorFlow [@martinabadi2015] to train a binary classifier (i.e. forest or non-forested). This methodology is increasingly being used to automate pixel-level classification in (color) digital photography data [@buscombe2018]. Training data were collected from the orthomosaic by randomly selecting 513 pixel square tiles from homogeneous forested or non-forested areas within the historical orthomosaic. Homogeneous tiles were combined in synthetic landscapes using a random gaussian field based binary mask (Figure 3). We generated 5000 synthetic landscapes for training, while 500 landscapes were generated for both the validation and the testing dataset. Source tiles did not repeat across datasets to limit overfitting. In order to limit stitch line misclassifications, along the seams of mosaicked images, we created synthetic landscapes with different forest tiles to mimick forest texture transitions. We applied this technique to 10% of the generated synthetic landscapes. The CNN model was trained for 100 epochs on a graphics processing unit (GPU) maximizing the Intersect-over-Union (IoU) using additional data augmentation. Data augmentation included random cropping to 320 pixel squares, random orientation, scaling, perspective, contrast and brightness shifts and image blurring. During final model evaluation we report the IoU of our out-of-sample test datasets.  The optimized model was used to classify the complete orthomosaic using a moving window approach with a step size of 110 pixels and a majority vote across overlapping areas to limit segmentation edge effects. We refer to Figure 4 for a synoptic overview of the full deep learning learning workflow.

More details on the optimizer and other metrics etc.

```{r echo=FALSE, out.width='100%', fig.cap="orthomosaic", fig.cap="An example synthetic landscape, combining homogeneous forest and non-forest images into a patchy landscape using random gaussian field based masks. The left panel shows a combined synthetic landscape, while the right panel shows the corresponding forest (black) and non-forest (white) labels."}
knitr::include_graphics('./figures/synthetic_landscape.png')
```

```{r echo=FALSE, out.width='100%', fig.cap="orthomosaic", fig.cap="A diagram of the deep learning workflow followed in training a binary forest / non-forest cover convolutional neural net U-Net model to generate our forest cover map."}
knitr::include_graphics('./figures/cnn_diagram.pdf')
```

To map long term land-use and land-cover change in the Yangambi region we used the contemporary Global Forest Change version 1.6 data (GFC, tile 10N-020E) [@hansen2013]. Using this data we calculated the latest state of the forest with respect to the conditions at the start of 1958, 60 years earlier. In our analysis we only included forested pixels which recorded no loss throughout the whole 2000 - 2018 period. As the resolution of the historical forest classification exceeds that of the GFC map we downsampled our historical forest cover data to 30m GFC resolution. We masked out all water bodies using the Global Forest Change survey data mask layer, and limited the analysis to the right bank of the Congo river. Where the datasets overlap we provide summary statistics of deforestation, reforestation and afforestation.

### Landscape fragmentation & Above Ground Carbon estimates

To quantify changes in the structure of forest cover and its disturbances we used spatial pattern analysis landscape fragmentation metrics [@dale1999]. We report the ratio of edge to area and the fractal index to quantify landscape complexity of forest disturbances. Statistics were calculated for all forest disturbance patches larger than 1 ha using the R package landscapemetrics [@hesselbarth2019]. We provide mean and standard deviation on edge, area, their ratio and fractal dimension for both the historical and contemporary @hansen2013 forest cover maps.

We estimated above ground carbon (AGC) losses and gains over time using plot based averages of recent inventory data at permanent sampling plots in the area (Figure 2). We refer to @Kearsley2013 for the survey method and allometric relations used to scale the survey data. Unlike standard square 1 ha plots edge plots were set back 200m from forest edges and were 50x200m, with the 50m side of the plot along the forest edge and continuing 200m into the forest (Appendix Table 1). We further confirmed that forest edge plots did not show a significantly different AGC compared to those of non-edged / mixed forest plots (Mann Whitney U test, p < 0.05). Thus it was not necessary to explicitly quantify changes in AGC caused by edge effects. Moreover, we used the mean values of the mixed forest as representative for potential AGC losses. Despite the challenges inherent in quantifying AGC for forest edges we mapped the total extent of the edges in the contemporary landscape. To align our landscape analysis with exploratory analysis of the survey data we used a buffer of 200m to estimate the extent of forest edges and patches, up to the location of forest edge plots.

Surveys of old plantations show a large variation in AGC, depending on age and the crop type. For example, the AGC values varied from 168.67 to 86.55 g Mg C ha^-1^, for *Hevea brasiliensis* (rubber tree) and *Elaeis guineensis* (oil palm) plots respectively (@bustillo2018, personal communications). These higher values are in line with the mixed AGC estimates in the area, while the palm plantations resemble old-regrowth values (81.8 Mg C ha^-1^, see Table A1). We therefore use both the estimates of old-regrowth and mixed forest to estimate AGC for regrowth. We did not have sufficient data to account for individual changes in AGC across plantations.

## Canopy structure & FOTO texture analysis

We compared the structure of the canopy both visually and using Fourier Transform Textural Ordination (FOTO, @couteron2002). Fourier Transform Textural Ordination uses a principal component analysis (PCA) on radially averaged 2D Fourier spectra to characterize canopy (image) texture.  The FOTO technique was first described by @couteron2002 to quantify canopy stucture in relation to biomass and biodiversity, and can be used across multiple scenes using normalization [@Barbier2010].

We used an area of 400x400m (16 ha) around contemporary permanent sampling plot locations to quantify the general state of the canopy in both historical and contemporary remote sensing data. Contemporary data were resampled to the resolution of the historical data using a nearest neighbour interpolation for comparison. We used the first acquisition of a pan-chromatic Geo-Eye 1 stereo pair (GeoEye, 2012, Thornton, Colorado, U.S.A., order 737537, 2011-11-11 8:55h GMT or 9:55h local time). For permanent sampling plot locations in both historical and contemporary data a rectangular buffer of 200m around the centre of all plots was used to create small image subsets. All image subsets were processed using the R based FOTO implementation and package [@hufkens2019a] using global normalization, ensuring intercomparability of texture metrics across time and space [@Barbier2010]. We used a FOTO (moving) window size of 187 pixels (or ~150m), as described in @solorzano2018 and @Barbier2010, to ensure that multiple individual canopies could be included in the analysis. A buffer with a radius of 50m around the center of all permanent sampling plots was used to determine the average value of the first principle component (PC), explaining the majority of the variance in canopy texture. Where both the Geo-Eye data and the orthomosaic image overlapped we processed the intersecting region to explore a wall-to-wall comparison between past and current canopy texture metrics.

We used the standard FOTO methodology with fixed zones, instead of the moving window approach. The window size was set to the same size (187 pixels or ~150m) as used in the moving window analysis above. To account for illumination differences between the two scenes we applied histogram matching. No global normalization was applied, as the scene was processed as a whole. PC values from this analysis for all permanent sampling plots in both image scences were extracted using a buffer with a radius of 50m.

For both site based and scene analysis we correlate PC values with permanent sample plot inventory data such as stem density, above ground biomass and tree species richness. Additional comparisons are made between contemporary Geo-Eye data and the historical orthomosaic derived PC values. Due to the few available permanent sampling plots in both scenes we used a non-parametric paired signed rank (Wilcoxon) test to determine differences between the PC values of the Geo-Eye and historical orthomosaic image scenes across mono-dominant and mixed forest types. In all analysis, mono-dominant site 4 was removed from the analysis due to cloud contamination.

# Results

Our analysis provides a first spatially explicity historical composite of aerial survey images mapping LULCC within the Congo Basin. The use of high resolution historical images combined with structure from motion image processing techniques allowed us to mosaic old imagery across a large extent. The final orthomosaic composition of the Yangambi region resulted in an image scene covering approximately 733 million pixels across ~828 km^2^ (~23x36 km, Figure 2). The overall accuracy of the structure from motion orthomosaic composition was 0.88 m/pixel using the sparse cloud DEM for corrections at 45.8 m/pixel. The resulting georeferenced scene had a spatial accuracy of approximately 23m. Further georeferencing outside the SfM workflow reduced the mean error at the ground control points to 5.3 $\pm$ 4.9 px (~4.7 $\pm$ 4.3 m), with a median error of 2.9 px (2.6m). The orthomosaic served as input for all subsequent LULCC analysis with all derived maps provided with the manuscript repository (see data & code availability statements below).

```{r echo=FALSE, out.width='100%', fig.cap="orthomosaic", fig.cap="Overview of the final orthomosaic of the greater Yangambi region (A), a detailed inset (B) and the derived land use land cover change map displayed as the difference between a manual classification of the orthomosaic and a recent map by Hansen et al. 2013 (C) and the corresponding land cover map for inset B (D)."}
knitr::include_graphics('./figures/orthomosaic_maps.png')
```

## Land-use and Land-Cover Change & Above Ground Carbon

Our CNN deep learning classifier reached an Intersection-over-Union (IoU) accuracy of 97% on the detection of disturbed forest in out-of-sample test data. Scaling our classifier to the whole historical orthomosaic we detected 162 km^2^ (or ~20% of the scene) of disturbed forests. A large fraction of the disturbed area was restored in the period between the two acquisitions. In total, 99 km^2^, or little over half of the affected forest was restored (Figure 5C-D, dark blue). Recent deforested areas, as registered through satellite remote sensing, approximate 88 km^2^ (Table1, Figure 5C-D / light green).

Recent deforestation follows a distincly different pattern compared to historical patterns. Historical deforestation showed a classical fishbone pattern for forest clearing with very sharp edges, while current patterns are patchy and ad-hoc (figure 5D, dark blue and green colours respectively). These differences are reflected in the analysis of landscape metrics of forest loss. Between the historical and contemporary LULCC maps we see an increase in small disturbances, as indicated by the decreasing area of the mean patch size, down to 37 ha from 120 ha historically, while perimeter lengths remained approximately the same, with 8653m and 9902m for historical and contemporary maps respectively (Table 2). This shift in perimeter area ratio led to a similar change in the fractal index, increasing in value to 1.13 $\pm$ 0.07 from 1.10 $\pm$ 0.05 over time. Values closer to a fractal index of 2 suggest a more complex (fragmented) landscape.

A comparison of forest edge plots with mixed forest plots showed no signficant difference in AGC, or other reported values such as species richness, basal area or stem density (Mann Whitney U test, p < 0.05). Edge influence did not extend beyond 200 m from a forest edge, but still represented an area of 132 km^2^.

Changes in both land-use and land-cover led to concomitant changes in above ground carbon stocks. Recovery throughout the region was characterized for patches of forest and plantations. Assuming high density stands, based on previous work, this could amount to a carbon gains of 1592 Gg C, offsetting more recent losses of approximately 1408 Gg C. On the other hand, at the low end, if we assume a lower carbon density of 81.8 Mg C ha^-1^ this would result in a total carbon gain of 811 Gg C. Using our approach results indicate that overall deforestation around Yangambi has resulted in a loss of ~2416 Gg C in AGC stocks.

```{r echo = FALSE, eval = TRUE}
change_stats %>%
  kable(caption="Land use land cover change statistics of forest cover around Yangambi in the central Congo Basin. The data evaluates a difference between a historical (1958) aerial photography based survey and the Hansen et al. 2013 based satellite remote sensing data. Spatial coverage statistics are provided in square kilometers (km) and hectares (ha), rounded to the nearest integer as well as Above Ground Carbon (AGC) scaled using recent survey measurements.",
       booktabs = TRUE,
       digits = 3,
       row.names = TRUE,
       escape = FALSE,
       col.names = c("$km\\textsuperscript{2}$",
                     "ha",
                     "(Gg C $ha\\textsuperscript{-1}$")
         ) %>%
    add_header_above(c(" " = 3,"AGC" = 1)) %>%
    kable_styling(latex_options = c("hold_position"))
```

```{r echo = FALSE, eval = TRUE}
patch_metrics %>%
  kable(caption="Landscape metrics for historical and contemporary deforestation patterns. We report patch perimeter and area, their ratio and fractal dimension. Values are reportedpat as mean $\\pm$ standard deviation, across all deforestation patches.",
       booktabs = TRUE,
       row.names = TRUE,
       escape = FALSE,
              col.names = c("perimeter (m)",
                     "area (ha)",
                     "ratio ($m\\textsuperscript{-1}$)",
                     "fractal dimension")
         ) %>%
    kable_styling(latex_options = c("hold_position"))
```

## Canopy structure & FOTO texture analysis

Visual interpretation of the scenes provide evidence that most locations do not change dramatically with respect to canopy composition, except for the large areas of disturbances in contemporary fallow or young-regrowth plots. One marked difference is noted in the mono-dominant site 6 (Appendix Table 1). Here, the current mono-dominant *Brachystegia laurenttii* is a recent development, changing the canopy structure visibly during the last half century (Figure 6). The previous varied canopy structure gave way to a more dense and uniform canopy. This is reflected in a change of the FOTO PC value from 0.19 historically to its current value of 0.54. This historical value is similar to the mean of contemporary mono-dominant stands with PC averaging 0.34 $\pm$ 0.1, and is only slightly higher than historical values for a mixed forest (0.18 $\pm$ 0.08). The reverse pattern is seen in the contemporary PC values. Here, the value of 0.54 exceeds those of most mono-dominant stands (0.35 $\pm$ 0.08), and is even further removed from the values noted for mixed forests (0.12 $\pm$ 0.03).

```{r echo=FALSE, out.width='100%', fig.cap="Visual comparison between a historical (A) and contemporary (B) permanent sampling plot. The site is currently listed as a mono-dominant Brachystegia laurenttii stand. Note the structural differences with a 'courser' canopy structure in the historical image compared to the more closed contemporary stand."}
knitr::include_graphics('./figures/visual_comparison_psp.png')
```

Using only small subsets around existing permanent sampling plots we show distinct differences between forest types, with PC values in both historical and contemporary imagery markedly higher for the mono-dominant forest types compared to all others (Appendix Figure 3). Provided that the young-regrowth and fallow permanent sampling plots have seen recent disturbance the Wilcoxon signed rank test on the mixed and mono-dominant plots between the historical and contemporary PC values did not show a significant difference (p > 0.05). Similarly, no significant different using PC values extracted from the whole scene analysis was noted (p > 0.05). Any relationships between contemporary Geo-Eye data and permanent sampling plot measurements of Above Ground Biomass, stem density and species richness were non-significant (p > 0.05, Appendix Figures 4-6).

Furthermore, visual inspection of the scene wide analysis suggests historical scences do not show landscape wide canopy features (Figure 7 A-B), unlike the contemporary scene (Figure 7 C-D). Where the FOTO algorithm picks up landscape features such as changes in texture across the contemporary Geo-Eye scene (e.g. the river valley as a diagonal line in Figure 7D), however, no corresponding landscape patterns are found by the FOTO algorithm in the historical orthomosaic.

```{r echo=FALSE, out.width='100%', fig.cap="RGB visualizations of the first three principal components of scene wide FOTO texture analysis of historical and current (Geo-Eye) imagery. Current permanent sampling plots of mono-dominant, mixed, fallow and young (edge) forest plots are marked with open triangles, open circles, open squares and crosses, respectively."}
knitr::include_graphics('./figures/foto_maps.png')
```

# Discussion

Finely grained spatial data sources, such as remote sensing imagery, are rare before the satellite era (<1970). This lack of data limits our understanding of how forest structure has varied over longer time periods in especially remote areas. Long term assessment can be extended by using large inventories of historical aerial survey data [@kadmon1999; @song2015; @nita2018]. However, across the Congo Basin this data is rare, or difficult to access. Despite the difficulties in assessing hidden disturbances caused by deforestation, poorly recorded official assessments or simply those invisible to standard remote sensing techniques [@peres2006], remote sensing generally remains the best way to map and quantify LULCC [@houghton2012]. Despite these challenges in recovery and validation we present a first analysis of LULCC using historical aerial survey data for the central Congo Basin. We used a combination of Structure-from-Motion and a convolutional neural net deep learning approach to map deforestation across a large extent, providing a long term view of LULCC in the central Congo Basin

## Data recovery challenges

In our study the archive data recovered was limited to contact prints and therefore did not represent the true resolution of the original negative. In addition, analogue photography clearly produces a distinct softness compared to digital imagery (Figure 4). Despite favourable nadir image acquistions [@verhoeven2017] image softness combined with illumination effects between flight paths, and the self-similar nature of vast canopy expanses [@park2019; @simini2010; @sole1995], limited our ability to provide wall-to-wall coverage of the entire dataset containing 334 images. Few man made features in the scenes also made georeferencing challenging. Although the village of Yangambi provided a range of buildings as (hard-edge) references, other areas within the central Congo Basin might have fewer permanent structures and would require the use of soft-edged landscape features (e.g. trees, river outflows). Research has shown that soft-edged features can help georeference scenes even when containing few man-made features [@hughes2006], however accuracy becomes compromised. Our two step georeferencing approach resulted in a referencing accuracy of ~4.7 $\pm$ 4.3 m across reference points. However, it shoud be noted that referencing accuracy of the final scene is less constrained toward the edges of the scene.

## Land-use, Land-Cover change & Above Ground Carbon

When classifying our orthomosaic into forest and non-forest states we favoured a deep learning supervised classification using a Convolutional Neural Network over manual segmentation to guarantee an "apples-to-apples" comparison between the historical and the contemporary GFC forest cover maps. We acknowledge that both classifications use different features, i.e. spectral or spatial data, but attain a similarly high accuracy of 97% and 99% [@hansen2013], for the GFC and our CNN map respectively. Despite a number of challenges we were able to map and quantify a vast continuous area of tropical forest, with limited manual annotation requirements.

Our analysis shows that the majority of deforestation around Yangambi happened toward the late 1950's (162 km^2^). Considerable regrowth has occured since the aerial survey was executed (99 km^2^), and socio-economic instability prevented further large scale forest exploitation. In particular, many plantations have reached maturity and forest has re-established in previously cleared or disturbed areas. The majority of this forest recovery takes the form of isolated patches of forest but is offset by further deforestation of previously untouched forest. Generally, the function and structure of forests can be influenced by forest edges that are located up to 1km away however most effects are pronounced within the first 300m from the edge [@gascon2000]. Our analysis of edge effects on AGC has shown that the influence is negligible 200m away from the edge. @phillips2006 have shown similar weak responses to edge effects in the Amazon forest. Due to a lack of data on the extent (depth) of edge effects and their influence on AGC beyond 200m we did not include any estimates of carbon loss or gain within these zones. However, it must be stated that the influence of edges throughout the landscape was not marginal. as these areas would account for 132 km^2^. Thus edges could have a substantial negative [@brinck2017] or positive [@reinmann2017] influence on AGC. Similarly, uncertainties in how to explicitly correct for plantations in the landscape present a further challenge. Thus although our estimates are only indicative they do underscore the important influence of landscape structure when carbon accounting. However, our findings do not indicate that deforestation in Congo basin is declining, on the contrary.

Over the past half century there has been a clear shift in land use in Yangambi (Figure 5). Land use has shifted away from for example a regular fishbone deforestation pattern that emerges when (large scale) agricultural interests dominate the landscape [@arima2015], to a more fragmented landscape (Figure 5D). The latter former? is consistent with historical land management at the time of the aerial survey [@bustillo2018]. These regular patterns have since been reversed because of a decrease in large scale intensive agriculture, replaced by ad-hoc small scale subsistence farming with large perimeter to area relationships (i.e. ragged edges). Consequently, edge effects in the current landscape are far more pronounced than in the historical scene.

Visual inspection of the images also suggests that reforestation within the historically cleared areas and experimental plots is not necessarily limited to areas far removed from more densely populated areas. For example, large reforested areas exist close to the Congo stream and Yangambi village itself (Fig. 5). Here, regional political components, such as land leases and large scale ownership could have played a role in safeguarding some of these areas for rewilding or sustainable management [@arima2014; @larson2011]. Despite widespread anthropogenic influences throughout the tropics [@lewis2015] the retention of these forested areas show the potential of explicit or implicit protective policy measures (e.g. INERA concessions, @bustillo2018) on a multi-decadal time scale. Forest regrowth in non-continuous areas within Yangambi could increase landscape connectivity and help increase biodiversity [@vandeperre2018].

Our analysis therefore provides an opportunity to highlight and study those regions that have previously suffered confirmed long-term disturbances, and those that have been restored since. Assessments of old plantations and recovering clear-cut forests can serve as a guide to help estimate carbon storage capacity and forest recovery rates in managed and unmanaged conditions [@gourlet-fleury2013; @sader1988; @achard2014] over the mid- to long-term, in support of rewilding and general forest restoration [@arima2014; @larson2011; @vandeperre2018]. In addition, mapping long-term edge effects can further support research into issues such as receding forest edges [@gascon2000].

## Canopy structure & FOTO texture analysis

Finally the FOTO technique used to quantify relationships between canopy structure and forest characteristics rendered no valuable insights of either the historical orthomosaic or recent Geo-Eye scene. Similarly weak correlations were found previously by @solorzano2018. In contrast site based texture metric statistics did show correspondence between historical and contemporary satellite imagery. However, none of them were either consistent or significant. Although visual interpretation shows distinctly different canopy structures (Figure 5) the differences in how resolution is defined and issues related to image quality prevented us from quantifying these further. Unlike large scale studies by @Ploton2012 we could not scale this technique to historical data. We advise that future valorisation efforts should preferentially work from foto negatives (if available) to ensure optimal data quality in resolution, contrast and overall sharpness.

# Conclusion

Given the impact of tropical forest disturbances on atmospheric CO_2 emissions, biodiversity and ecosystem productivity accurate long term reporting on land-cover and land-use change especially in the pre-satellite era is an imperative. Our analysis of historical aerial survey images of the Central Congo Basin provides a window into the state of the forest at the start of the anthropocene. Efforts to quantify canopy texture and their link to carbon storage had limited to no success.We have shown that historical aerial survey data can be used to quantify land-use and land-cover change using a highly automated workflow to quantify the structure and extent of forest cover change that can help assess the impact of fragmentation on above ground carbon stocks. We also show a shift from previously highly structured industrial deforestation of large areas, to discrete smallholder clearing for farming, increasing landscape fragmentation but also opportuntiy for substantial regrowth. Efforts to quantify canopy texture and their link to carbon storage had limited to no success. As such, our analysis provides insights into the state of rarely studied tropical forests and the rate at which deforestation and reforestation has taken place over a multi-decadal scale in the central Congo basin providing a useful historical context to land-cover and land-use change spatial patterns for interpreting past and ongoing field research in the area.

# Additional Information and Declarations
## Data availability

Hufkens et al. (2019). A curated dataset of aerial survey images over the central Congo Basin, 1958. Zenodo: doi.org/10.5281/zenodo.3547767. All data not included in the latter repository can be found bundled with the analysis code as listed below. Proprietary datasets (i.e. Geo-Eye data) are not shared, but purchase order numbers allow for acquisition of these datasets to ensure reproducibility. The @hansen2013 image data is freely available from [the GFC project website](https://earthenginepartners.appspot.com/science-2013-global-forest/download_v1.6.html).

## Code availability

All analysis code is available as an R / python [@rcoreteam2019] projects (https://khufkens.github.io/orthodrc & https://khufkens.github.io/orthodrc_cnn/). The analysis relied heavily on the 'raster' [@hijmans2019], 'RStoolbox' [@leutner2019], and 'landscapemetrics' [@hesselbarth2019] packages, while post-processing and plotting was facilitated by the 'tidyverse' ecosystem [@wickham2017], 'ggthemes' [@arnold2019], 'scales' [@wickham2018] and 'cowplot' [@wilke2019]. Additional plotting elements were formatted or provided by 'sf' [@pebesma2018] and 'rnaturalearth' [@south2017] packages, respectively. I'm grateful for the contributions to the scientific community by the developers of these packages.