### 2019-12-27 manuscript/main.tex

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[remote sensing,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

\usepackage{longtable}

% see https://stackoverflow.com/a/47122900

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, neuroglia, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, polysaccharides, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, stats, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, vibration, viruses, vision, water, wem
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2018}
\copyrightyear{2018}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Historical aerial surveys map long-term changes of forest cover and
structure in the central Congo Basin}

% Authors, for the paper (add full first names)
\Author{Koen Hufkens$^{1,2,*}$\href{https://orcid.org/0000-0002-5070-8109}{\orcidicon}, Thalès de Haulleville$^{3}$, Elizabeth Kearsley$^{1}$, Kim Jacobsen$^{2,3}$, Hans Beeckman$^{3}$, Piet Stoffelen$^{4}$, Filip Vandelook$^{4}$, Sofie Meeus$^{4}$, Michael Amara$^{5}$, Leen Van Hirtum$^{5}$, Jan Van den Bulcke$^{1}$, Hans Verbeeck$^{1}$, Lisa Wingate$^{2}$}

% Authors, for metadata in PDF
\AuthorNames{Koen Hufkens, Thalès de Haulleville, Elizabeth Kearsley, Kim Jacobsen, Hans Beeckman, Piet Stoffelen, Filip Vandelook, Sofie Meeus, Michael Amara, Leen Van Hirtum, Jan Van den Bulcke, Hans Verbeeck, Lisa Wingate}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Ghent University, Ghent, Belgium; \\
$^{2}$ \quad INRA, UMR ISPA, Villenave d'Ornon, France; \\
$^{3}$ \quad Royal Museum for Central Africa, Tervuren, Belgium; \\
$^{4}$ \quad Botanic Garden Meise, Meise, Belgium; \\
$^{5}$ \quad National Archives of Belgium, Brussels, Belgium; \\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:koen.hufkens@gmail.com}{\nolinkurl{koen.hufkens@gmail.com}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Land Use and Land Cover change (LULCC) of African rainforest contribute
to global carbon emissions. Yet, most historical estimates of LULCC and
their carbon emissions rely on non-spatially explicity data in the
pre-satellite era (\textless{} 1972). Past studies have expanded this
pre-satellite time frame with declassified satellite surveillance data
or aerial surveys, but none cover the Congo Basin. Here, we use
historical aerial survey photos to map the extent and structure of LULCC
around Yangambi, central Congo Basin, in 1958. Our study leveraged
Structure-from-Motion to generate a large orthomosaic covering 828
km\textsuperscript{2}, counting 733 million pixels, at a sub meter
resolution and geo-referenced to \textasciitilde{}4.7 \(\pm\) 4.3 m.
Primary forest in the orthomosaic was classified with a Convoluted
Neural Network Deep Learning approach. We used these data to quantify
LULCC, landscape and canopy complexity and scale above ground biomass
between historical and contemporary condition. We show a shift from
historical highly structured industrial deforestation in 1958 (162
km\textsuperscript{2}), to contemporary smallholder farming clearing (88
km\textsuperscript{2}), increasing landscape fragmentation and forest
edges. Efforts to quantify canopy texture and their link to carbon
storage had limited to no success. Our analysis provides an insight in
key spatial and temporal patterns of deforestation and reforestation at
a multi-decadal scale and provide a historical context to land-cover and
land-use change spatial patterns for past and ongoing field research in
the area.}

% Keywords
\keyword{Aerial survey, data recovery, CNN, Deep Learning, SfM, Congo Basin}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{wrapfig} \usepackage{graphics} \usepackage{setspace}

\doublespacing

\usepackage{gensymb} \usepackage{booktabs}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The functioning of a tropical forest, and its capacity to sequester
carbon, is heavily influenced by anthropogenic disturbances such as
selective logging, clear felling, or slash and burn practices, all
increasing forest fragmentation. This increased pace in the loss of
rainforest cover \citep{butsic2015} is predicted to negatively impact
the function of these tropical forests, and their future capacity to
sequester carbon, as anthropogenic disturbances such as selective
logging, clear felling, or slash and burn practices collectively
contribute to increasing forest fragmentation.

These forest fragments and their edges contain fewer large trees
\citep{laurance2000}, in part due to their hotter and brighter
microclimates \citep{magnago2015}. Similarly, species composition and
biodiversity \citep{poorter2006} are negatively affected. Spatial
disturbances also have an explicit temporal component, where the time
since disturbance affects the strength of the effects
\citep{didham1999}. Deforestation and gap formation also influence the
overall canopy structure, where pioneer species grow taller and with
narrower crowns \citep{poorter2006, peres2006} affecting the forest
light regime and carbon dynamics \citep{stark2012}. As such,
anthropogenic factors through land-use and land cover change (LULCC)
heavily affect forest structure and functioning
\citep{fauset2017, brinck2017, didham1999}. In the tropics the majority
of emissions originate from deforestation and forest degradation
\citep{hansen2013, houghton2012, tyukavina2015} contributing to 10-15\%
of the total global emissions \citep{vanderwerf2009} and concomitant
losses in biodiversity losses \citep{barlow2016, vandeperre2018} and how
LULCC is evolving over time. Although the African rainforest is the
second largest on Earth and covers \textasciitilde{}630 million ha,
representing up to 66 Pg of carbon storage \citep{Lewis2009b}, and
currently loses forest at an increasing pace \citep{butsic2015}. Yet,
the tropics is now becoming a large source of CO\_2 to the atmosphere as
deforestation and forest degradation
\citep{hansen2013, houghton2012, tyukavina2015} contribute up to 10-15\%
of the total global CO2 emissions \citep{vanderwerf2009}. Furthermore,
recent estimates show that 31\% of recent carbon emissions are caused by
edge effects alone \citep{brinck2017}. In this context, there are very
few long term forest inventory plots that presently exist in the central
Congo Basin \citep{Verbeeck2011a, lewis2013}. This further limits the
accuracy of current biomass estimates and challenges our understanding
of how forest structure and function changes in the short and long term
in these ecosystems \citep{Kearsley2013, Doetterl2015}.

Historical estimates of LULCC, and their carbon emissions, exist
\citep{houghton2012, ramankutty1999, defries2002, sader1988}, however
they generally rely on non-spatially explicit data in the pre-satellite
era (\textless{} 1972) \citep{houghton2010}. These analysis often use
FAO statistics, historical land-use sources, and/or population and
socio-economic data \citep{houghton2012}, as limited earlier spatial
data sources are available \citep{hansen2013, achard2014}. For example,
the \citet{houghton2010} long-term study used FAO and national censuses
alongside a bookkeeping model to estimate carbon emissions from LULCC.
Other studies use remote sensing data, but are limited in time to recent
decades \citep{defries2002, houghton2012}. These proxy based studies
have a limited view on long-term direct and indirect spatial effects of
deforestation. As such, both the spatial structure and the fragmentation
pattern of a forest combined with its life history has a profound effect
on carbon emssisions. However, in this context remote sensing
assessments provide more accurate spatial information to help estimate
carbon emissions \citep{mitchard2018}.

Past studies expanded knowledge during the pre-satellite time frame
through the use of now declassified satellite surveillance data to map
deforestation in the US, central Brazil, and post world war II Germany
\citep{song2015, nita2018}. In some cases these records could extend
back in time even further using historical aerial survey images. In
northern Europe especially these data are well known and used, for
example to map the extent of forest cover across \textasciitilde{}100
km\textsuperscript{2} regularly, and at times using automated workflows
\citep{giordano2018}. Survey data across the African continent is less
common or inaccessible or both. Some studies do exist, both
\citet{buitenwerf2012} and \citet{okeke2006} used aerial survey images
to map vegetation changes in South African savannas and the Adulam
Nature Reserve, respectively. Whilst \citet{frankl2015} used 20 aerial
survey images to map the Ethiopian highlands of 1935. High resolution
aerial or surveillance satellite images provide scientists tools to
monitor forest extent and structure over long time frames. Moreover,
canopy texture from aerial survey images has been successfully linked to
above ground biomass \citep{Ploton2012, Couteron2005, Barbier2010}. Thus
using texture based metrics obtained from (high resolution) remote
sensing data opens up possibilities to link historical aerial footage to
carbon density across large spatial extents \citep{Couteron2005}.

Given the impact of LULCC on the structure and functioning of tropical
forests, and their influence on both carbon dynamics
\citep{Kearsley2013} and biodiversity \citep{vandeperre2018}, accurate
long term reporting of historical forest cover in the pre-satellite era
is an imperative \citep{willcock2016}. In this study we use a
combination of historical aerial photography and contemporary remote
sensing data to map long-term changes in the extent and structure of the
tropical forest surrounding Yangambi, in the central Congo Basin at the
beginning of 1958 and the start of the antrophocene \citep{lewis2015}.
Our analysis provides a historical insight into important LULCC spatial
patterns in Yangambi, such as fragmentation and edge complexity. Our
analysis further contextualises the influence of changes in the forest's
life history on past and current research into Above Ground Carbon (AGC)
storage \citep{Kearsley2013} and biodiversity \citep{vandeperre2018} in
the central Congo Basin.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{historical-data-acquisition}{%
\subsection{Historical data
acquisition}\label{historical-data-acquisition}}

Data for the central Congo Basin region, surrounding Kisangani, were
collected in several flights during the dry season of 1958 and 1959
(from 8/01/1958 to 20/02/1958 and from 28/12/1958 to 9/01/1959
respectively, see Appendix Figure 1) to generate topographic maps of the
area, supervised by the ``Institut Géographique du Congo Belge'' in
Kinshasa (then Léopoldville). Black-and-White infrared images (0.4 - 0.9
\(\mu\)m) were gathered along flight paths running mostly from west to
east, between 9 - 11h local time. Along a flight path continuous images
were taken using a Wild Heerburgg RC5a (currently Leica Geosystems) with
an Aviogon lens assembly (114.83mm / f 5.6, with a 90\degree view angle)
resulting in square photo negative of 180 by 180 mm. Flights were flown
at an average absolute altitude of \textasciitilde{}5200 m above sea
level, covering roughly 18 530 km\textsuperscript{2} at an approximate
scale of 1/40 000. The use of the integrated autograph system ensured
timely acquisition of pictures with a precise overlap
(\textasciitilde{}1/3) between images. This large overlap between images
together with flight parameters would allow post-processing, using
stereographs, to create accurate topographic maps. Original data from
this campaign are stored in the
\href{https://www.africamuseum.be/en}{Royal Museum for Central Africa}
in Tervuren, Belgium (Figure 1).

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/print_examples} \caption{A box of historical aerial photographs (left) and a single aerial photograph (right) showing part of the Congo river. Note the meta-data provided in the margins of the image such as acquisition time and fleight height and the solar glare on the Congo river.}\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{site-selection}{%
\subsection{Site selection}\label{site-selection}}

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/flight_paths} \caption{Overview of the historical flight paths during aerial photo acquisition and ancillary data used in this study. The bounding box of the orthomosaic data presented in this study is shown as a rectangle (23x36 km). The outline of a recent high-resolution Geo-eye panchromatic image is shown as a dashed dark grey rectangle (10x10km). The location of various permanent sampling plots are shown as x, +, and open squares and triangles for the mixed, mono-dominant and edge plots respectively. The grey polygon delineates the current day Yangambi Man and Biosphere reserve. The inset, top right, situates the greater Yangabmi region with the DR Congo. The full flight plan and details are shown in Appendix Figure 1 and 2}\label{fig:unnamed-chunk-2}
\end{figure}

We prioritised flight paths and images that contained current day
permanent sampling plots, larger protected areas, and past agricultural
and forest research facilities (Figure 2). This selection provides a
comprehensive mapping of the Yangambi area and the life history of the
forest surrounding it. Thereafter, we selected flight paths 1 through 11
for digitization. From this larger dataset of 334 images we selected 74
survey images for orthomosaic compositing and further analysis. All the
selected images stem from the flight campaign made during January and
February of 1958. The area includes the Yangambi village, 20
contemporary permanent sampling plots \citep{Kearsley2013}, past and
present agricultural experimental plots \citep{bauters2015} and large
sections of the Yangambi
\href{http://www.unesco.org/new/en/natural-sciences/environment/ecological-sciences/biosphere-reserves/africa/democratic-republic-of-the-congo/yangambi/}{UNESCO
Man and Biosphere reserve} surrounding to the west and east of the
village. Although not formally mosaicked we provided a full dataset of
pre-processed images using the cropping and normalization routines
described below. The latter data was not used in subsequent LULCC
analysis, but has been archived and made available to the public
separately (see code \& data availability statement below).

\hypertarget{digitization-and-data-processing}{%
\subsection{Digitization and data
processing}\label{digitization-and-data-processing}}

All selected images, covering the Yangambi area, were contact prints as
original negatives of the prints were not available. Images were scanned
at a resolution exceeding their original resolution (or grain) at the
maximal physical resolution of an Epson A3 flatbed scanner (i.e.~2400
dpi or 160MP per image) and saved as lossless tiff images. Data were
normalized using contrast limited histogram equalization
\citep{zuiderveld1994} with a window size of 32 and a clip limit of 1.5.
Fiduciary marks were used to rectify and downsample the images into
square 7700x7700 pixel images (\textasciitilde{}1200 dpi, 81MP). This
resulted in a dataset with digital images at a resolution that remained
above the visible grain of the photographs, whilst the reduced image
size facilitated easier file handling and processing speed.

Data was processed into a georeferenced orthomosaic using a Structure
from motion (SfM, \citet{ullman1979}) approach implemented in
\href{http://www.agisoft.com}{Agisoft Metashape} version 1.5.2 (Agisoft
LLC, St.~Petersburg, Russia). An orthomosaic corrects remote sensing
data to represent a perfectly downward looking image, free from
perspective distortions due to topography and camera tilt. Using the SfM
technique features, areas in images with a large degree of similarity,
are matched across various images to reconstruct a three dimensional
scene (topography) from two-dimensional image sequences. During the SfM
analysis we masked clouds, glare or large water bodies such as the Congo
river.

We calculated the orthomosaic using a low resolution point cloud and
digital elevation map (DEM). Additional ground control points were
provided to assist in the referencing of image and constrain the
optimization routine used in the SfM algorithm. Ground control points
consisted of rooftop edges of permanent structures which could be
verified in both old and new aerial imagery (i.e.~ESRI World Imagery).
Although clouds were removed during the SfM routine we did not mask all
clouds in the final orthomosaic to maximize forest coverage. The final
scene was cropped, to provide consistent wall to wall coverage of the
reconstructed scene. The orthomosaic was exported as a geotiff for
further georeferencing in QGIS \citep{qgisdevelopmentteam2018} using the
georeferencer plugin (version 3.1.9) and additional ESRI World Imagery
high resolution reference data. We used 3\textsuperscript{th} degree
polynomial and 16 ground control points to correct the final image.
Ground control points, raw image data and final processed image are
provided in addition to measures of uncertainty such as mean root mean
squared (RMSE), mean and median error across all ground control points.
All subsequent analysis are executed on the final geo-referenced
orthomosaic or subsets of it.

\hypertarget{land-use-and-land-cover-change}{%
\subsection{Land-Use and Land-Cover
Change}\label{land-use-and-land-cover-change}}

\hypertarget{classifying-forest-cover}{%
\subsubsection{Classifying forest
cover}\label{classifying-forest-cover}}

We automatically delineated all natural forest in the historical data,
thus excluding tree plantations, thinned or deteriorated forest stands,
fields and buildings. We used the Unet Convoluted Neural Net (CNN,
\citet{ronneberger2015}) architecture implemented in Keras
\citep{chollet2015} with an efficientnetb3 backbone
\citep{yakubovskiy2019} running on TensorFlow \citep{martinabadi2015} to
train a binary classifier (i.e.~forest or non-forested). This
methodology is increasingly being used to automate pixel-level
classification in (color) digital photography data \citep{buscombe2018}.
Training data were collected from the orthomosaic by randomly selecting
513 pixel square tiles from homogeneous forested or non-forested areas
within the historical orthomosaic. Homogeneous tiles were combined in
synthetic landscapes using a random gaussian field based binary mask
(Figure 3). We generated 5000 synthetic landscapes for training, while
500 landscapes were generated for both the validation and the testing
dataset. Source tiles did not repeat across datasets to limit
overfitting. In order to limit stitch line misclassifications, along the
seams of mosaicked images, we created synthetic landscapes with
different forest tiles to mimick forest texture transitions. We applied
this technique to 10\% of the generated synthetic landscapes. The CNN
model was trained for 100 epochs on a graphics processing unit (GPU)
maximizing the Intersect-over-Union (IoU) using additional data
augmentation. Data augmentation included random cropping to 320 pixel
squares, random orientation, scaling, perspective, contrast and
brightness shifts and image blurring. During final model evaluation we
report the IoU of our out-of-sample test datasets. The optimized model
was used to classify the complete orthomosaic using a moving window
approach with a step size of 110 pixels and a majority vote across
overlapping areas to limit segmentation edge effects. We refer to Figure
4 for a synoptic overview of the full deep learning learning workflow.

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/synthetic_landscape} \caption{An example synthetic landscape, combining homogeneous forest and non-forest images into a patchy landscape using random gaussian field based masks. The left panel shows a combined synthetic landscape, while the right panel shows the corresponding forest (black) and non-forest (white) labels.}\label{fig:unnamed-chunk-3}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/cnn_diagram} \caption{A diagram of the deep learning workflow followed in training a binary forest / non-forest cover convolutional neural net U-Net model to generate our forest cover map.}\label{fig:unnamed-chunk-4}
\end{figure}

To map long term land-use and land-cover change in the Yangambi region
we used the contemporary Global Forest Change version 1.6 data (GFC,
tile 10N-020E) \citep{hansen2013}. Using this data we calculated the
latest state of the forest with respect to the conditions at the start
of 1958, 60 years earlier. In our analysis we only included forested
pixels which recorded no loss throughout the whole 2000 - 2018 period.
As the resolution of the historical forest classification exceeds that
of the GFC map we downsampled our historical forest cover data to 30m
GFC resolution. We masked out all water bodies using the Global Forest
Change survey data mask layer, and limited the analysis to the right
bank of the Congo river. Where the datasets overlap we provide summary
statistics of deforestation, reforestation and afforestation.

\hypertarget{landscape-fragmentation-above-ground-carbon-estimates}{%
\subsubsection{Landscape fragmentation \& Above Ground Carbon
estimates}\label{landscape-fragmentation-above-ground-carbon-estimates}}

To quantify changes in the structure of forest cover and its
disturbances we used spatial pattern analysis landscape fragmentation
metrics \citep{dale1999}. We report the ratio of edge to area and the
fractal index to quantify landscape complexity of forest disturbances.
Statistics were calculated for all forest disturbance patches larger
than 1 ha using the R package landscapemetrics \citep{hesselbarth2019}.
We provide mean and standard deviation on edge, area, their ratio and
fractal dimension for both the historical and contemporary
\citet{hansen2013} forest cover maps.

We estimated above ground carbon (AGC) losses and gains over time using
plot based averages of recent inventory data at permanent sampling plots
in the area (Figure 2). We refer to \citet{Kearsley2013} for the survey
method and allometric relations used to scale the survey data. Unlike
standard square 1 ha plots edge plots were set back 200m from forest
edges and were 50x200m, with the 50m side of the plot along the forest
edge and continuing 200m into the forest (Appendix Table 1). We further
confirmed that forest edge plots did not show a significantly different
AGC compared to those of non-edged / mixed forest plots (Mann Whitney U
test, p \textless{} 0.05). Thus it was not necessary to explicitly
quantify changes in AGC caused by edge effects. Moreover, we used the
mean values of the mixed forest as representative for potential AGC
losses. Despite the challenges inherent in quantifying AGC for forest
edges we mapped the total extent of the edges in the contemporary
landscape. To align our landscape analysis with exploratory analysis of
the survey data we used a buffer of 200m to estimate the extent of
forest edges and patches, up to the location of forest edge plots.

Surveys of old plantations show a large variation in AGC, depending on
age and the crop type. For example, the AGC values varied from 168.67 to
86.55 g Mg C ha\textsuperscript{-1}, for \emph{Hevea brasiliensis}
(rubber tree) and \emph{Elaeis guineensis} (oil palm) plots respectively
(\citet{bustillo2018}, personal communications). These higher values are
in line with the mixed AGC estimates in the area, while the palm
plantations resemble old-regrowth values (81.8 Mg C
ha\textsuperscript{-1}, see Table A1). We therefore use both the
estimates of old-regrowth and mixed forest to estimate AGC for regrowth.
We did not have sufficient data to account for individual changes in AGC
across plantations.

\hypertarget{canopy-structure-foto-texture-analysis}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis}}

We compared the structure of the canopy both visually and using Fourier
Transform Textural Ordination (FOTO, \citet{couteron2002}). Fourier
Transform Textural Ordination uses a principal component analysis (PCA)
on radially averaged 2D Fourier spectra to characterize canopy (image)
texture. The FOTO technique was first described by \citet{couteron2002}
to quantify canopy stucture in relation to biomass and biodiversity, and
can be used across multiple scenes using normalization
\citep{Barbier2010}.

We used an area of 400x400m (16 ha) around contemporary permanent
sampling plot locations to quantify the general state of the canopy in
both historical and contemporary remote sensing data. Contemporary data
were resampled to the resolution of the historical data using a nearest
neighbour interpolation for comparison. We used the first acquisition of
a pan-chromatic Geo-Eye 1 stereo pair (GeoEye, 2012, Thornton, Colorado,
U.S.A., order 737537, 2011-11-11 8:55h GMT or 9:55h local time). For
permanent sampling plot locations in both historical and contemporary
data a rectangular buffer of 200m around the centre of all plots was
used to create small image subsets. All image subsets were processed
using the R based FOTO implementation and package \citep{hufkens2019a}
using global normalization, ensuring intercomparability of texture
metrics across time and space \citep{Barbier2010}. We used a FOTO
(moving) window size of 187 pixels (or \textasciitilde{}150m), as
described in \citet{solorzano2018} and \citet{Barbier2010}, to ensure
that multiple individual canopies could be included in the analysis. A
buffer with a radius of 50m around the center of all permanent sampling
plots was used to determine the average value of the first principle
component (PC), explaining the majority of the variance in canopy
texture. Where both the Geo-Eye data and the orthomosaic image
overlapped we processed the intersecting region to explore a
wall-to-wall comparison between past and current canopy texture metrics.

We used the standard FOTO methodology with fixed zones, instead of the
moving window approach. The window size was set to the same size (187
pixels or \textasciitilde{}150m) as used in the moving window analysis
above. To account for illumination differences between the two scenes we
applied histogram matching. No global normalization was applied, as the
scene was processed as a whole. PC values from this analysis for all
permanent sampling plots in both image scences were extracted using a
buffer with a radius of 50m.

For both site based and scene analysis we correlate PC values with
permanent sample plot inventory data such as stem density, above ground
biomass and tree species richness. Additional comparisons are made
between contemporary Geo-Eye data and the historical orthomosaic derived
PC values. Due to the few available permanent sampling plots in both
scenes we used a non-parametric paired signed rank (Wilcoxon) test to
determine differences between the PC values of the Geo-Eye and
historical orthomosaic image scenes across mono-dominant and mixed
forest types. In all analysis, mono-dominant site 4 was removed from the
analysis due to cloud contamination.

\hypertarget{results}{%
\section{Results}\label{results}}

Our analysis provides a first spatially explicity historical composite
of aerial survey images mapping LULCC within the Congo Basin. The use of
high resolution historical images combined with structure from motion
image processing techniques allowed us to mosaic old imagery across a
large extent. The final orthomosaic composition of the Yangambi region
resulted in an image scene covering approximately 733 million pixels
across \textasciitilde{}828 km\textsuperscript{2}
(\textasciitilde{}23x36 km, Figure 2). The overall accuracy of the
structure from motion orthomosaic composition was 0.88 m/pixel using the
sparse cloud DEM for corrections at 45.8 m/pixel. The resulting
georeferenced scene had a spatial accuracy of approximately 23m. Further
georeferencing outside the SfM workflow reduced the mean error at the
ground control points to 5.3 \(\pm\) 4.9 px (\textasciitilde{}4.7
\(\pm\) 4.3 m), with a median error of 2.9 px (2.6m). The orthomosaic
served as input for all subsequent LULCC analysis with all derived maps
provided with the manuscript repository (see data \& code availability
statements below).

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/orthomosaic_maps} \caption{Overview of the final orthomosaic of the greater Yangambi region (A), a detailed inset (B) and the derived land use land cover change map displayed as the difference between a manual classification of the orthomosaic and a recent map by Hansen et al. 2013 (C) and the corresponding land cover map for inset B (D).}\label{fig:unnamed-chunk-5}
\end{figure}

\hypertarget{land-use-and-land-cover-change-above-ground-carbon}{%
\subsection{Land-use and Land-Cover Change \& Above Ground
Carbon}\label{land-use-and-land-cover-change-above-ground-carbon}}

Our CNN deep learning classifier reached an Intersection-over-Union
(IoU) accuracy of 97\% on the detection of disturbed forest in
out-of-sample test data. Scaling our classifier to the whole historical
orthomosaic we detected 162 km\textsuperscript{2} (or
\textasciitilde{}20\% of the scene) of disturbed forests. A large
fraction of the disturbed area was restored in the period between the
two acquisitions. In total, 99 km\textsuperscript{2}, or little over
half of the affected forest was restored (Figure 5C-D, dark blue).
Recent deforested areas, as registered through satellite remote sensing,
approximate 88 km\textsuperscript{2} (Table1, Figure 5C-D / light
green).

Recent deforestation follows a distincly different pattern compared to
historical patterns. Historical deforestation showed a classical
fishbone pattern for forest clearing with very sharp edges, while
current patterns are patchy and ad-hoc (figure 5D, dark blue and green
colours respectively). These differences are reflected in the analysis
of landscape metrics of forest loss. Between the historical and
contemporary LULCC maps we see an increase in small disturbances, as
indicated by the decreasing area of the mean patch size, down to 37 ha
from 120 ha historically, while perimeter lengths remained approximately
the same, with 8653m and 9902m for historical and contemporary maps
respectively (Table 2). This shift in perimeter area ratio led to a
similar change in the fractal index, increasing in value to 1.13 \(\pm\)
0.07 from 1.10 \(\pm\) 0.05 over time. Values closer to a fractal index
of 2 suggest a more complex (fragmented) landscape.

A comparison of forest edge plots with mixed forest plots showed no
signficant difference in AGC, or other reported values such as species
richness, basal area or stem density (Mann Whitney U test, p \textless{}
0.05). Edge influence did not extend beyond 200 m from a forest edge,
but still represented an area of 132 km\textsuperscript{2}.

Changes in both land-use and land-cover led to concomitant changes in
above ground carbon stocks. Recovery throughout the region was
characterized for patches of forest and plantations. Assuming high
density stands, based on previous work, this could amount to a carbon
gains of 1592 Gg C, offsetting more recent losses of approximately 1408
Gg C. On the other hand, at the low end, if we assume a lower carbon
density of 81.8 Mg C ha\textsuperscript{-1} this would result in a total
carbon gain of 811 Gg C. Using our approach results indicate that
overall deforestation around Yangambi has resulted in a loss of
\textasciitilde{}2416 Gg C in AGC stocks.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-6}Land use land cover change statistics of forest cover around Yangambi in the central Congo Basin. The data evaluates a difference between a historical (1958) aerial photography based survey and the Hansen et al. 2013 based satellite remote sensing data. Spatial coverage statistics are provided in square kilometers (km) and hectares (ha), rounded to the nearest integer as well as Above Ground Carbon (AGC) scaled using recent survey measurements.}
\centering
\begin{tabular}[t]{lrrl}
\toprule
\multicolumn{3}{c}{ } & \multicolumn{1}{c}{AGC} \\
\cmidrule(l{3pt}r{3pt}){4-4}
  & $km\textsuperscript{2}$ & ha & (Gg C $ha\textsuperscript{-1}$\\
\midrule
Forest & 685 & 68455 & -\\
-  of which edges & 132 & 13151 & -\\
Regrowth > 1958 loss & 99 & 9918 & 811 - 1592\\
Loss > 2000 & 88 & 8776 & 1408\\
Loss > 1958 (permanent) & 63 & 6282 & 1008\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-7}Landscape metrics for historical and contemporary deforestation patterns. We report patch perimeter and area, their ratio and fractal dimension. Values are reportedpat as mean $\pm$ standard deviation, across all deforestation patches.}
\centering
\begin{tabular}[t]{lllll}
\toprule
  & perimeter (m) & area (ha) & ratio ($m\textsuperscript{-1}$) & fractal dimension\\
\midrule
historical & 8653 $\pm$ 35676 & 120 $\pm$ 659 & 0.031 $\pm$ 0.014 & 1.099 $\pm$ 0.052\\
contemporary & 9902 $\pm$ 56975 & 37 $\pm$ 226 & 0.046 $\pm$ 0.014 & 1.13 $\pm$ 0.071\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{canopy-structure-foto-texture-analysis-1}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-1}}

Visual interpretation of the scenes provide evidence that most locations
do not change dramatically with respect to canopy composition, except
for the large areas of disturbances in contemporary fallow or
young-regrowth plots. One marked difference is noted in the
mono-dominant site 6 (Appendix Table 1). Here, the current mono-dominant
\emph{Brachystegia laurenttii} is a recent development, changing the
canopy structure visibly during the last half century (Figure 6). The
previous varied canopy structure gave way to a more dense and uniform
canopy. This is reflected in a change of the FOTO PC value from 0.19
historically to its current value of 0.54. This historical value is
similar to the mean of contemporary mono-dominant stands with PC
averaging 0.34 \(\pm\) 0.1, and is only slightly higher than historical
values for a mixed forest (0.18 \(\pm\) 0.08). The reverse pattern is
seen in the contemporary PC values. Here, the value of 0.54 exceeds
those of most mono-dominant stands (0.35 \(\pm\) 0.08), and is even
further removed from the values noted for mixed forests (0.12 \(\pm\)
0.03).

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/visual_comparison_psp} \caption{Visual comparison between a historical (A) and contemporary (B) permanent sampling plot. The site is currently listed as a mono-dominant Brachystegia laurenttii stand. Note the structural differences with a 'courser' canopy structure in the historical image compared to the more closed contemporary stand.}\label{fig:unnamed-chunk-8}
\end{figure}

Using only small subsets around existing permanent sampling plots we
show distinct differences between forest types, with PC values in both
historical and contemporary imagery markedly higher for the
mono-dominant forest types compared to all others (Appendix Figure 3).
Provided that the young-regrowth and fallow permanent sampling plots
have seen recent disturbance the Wilcoxon signed rank test on the mixed
and mono-dominant plots between the historical and contemporary PC
values did not show a significant difference (p \textgreater{} 0.05).
Similarly, no significant different using PC values extracted from the
whole scene analysis was noted (p \textgreater{} 0.05). Any
relationships between contemporary Geo-Eye data and permanent sampling
plot measurements of Above Ground Biomass, stem density and species
richness were non-significant (p \textgreater{} 0.05, Appendix Figures
4-6).

Furthermore, visual inspection of the scene wide analysis suggests
historical scences do not show landscape wide canopy features (Figure 7
A-B), unlike the contemporary scene (Figure 7 C-D). Where the FOTO
algorithm picks up landscape features such as changes in texture across
the contemporary Geo-Eye scene (e.g.~the river valley as a diagonal line
in Figure 7D), however, no corresponding landscape patterns are found by
the FOTO algorithm in the historical orthomosaic.

\begin{figure}
\includegraphics[width=1\linewidth]{./figures/foto_maps} \caption{RGB visualizations of the first three principal components of scene wide FOTO texture analysis of historical and current (Geo-Eye) imagery. Current permanent sampling plots of mono-dominant, mixed, fallow and young (edge) forest plots are marked with open triangles, open circles, open squares and crosses, respectively.}\label{fig:unnamed-chunk-9}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Finely grained spatial data sources, such as remote sensing imagery, are
rare before the satellite era (\textless{}1970). This lack of data
limits our understanding of how forest structure has varied over longer
time periods in especially remote areas. Long term assessment can be
extended by using large inventories of historical aerial survey data
\citep{kadmon1999, song2015, nita2018}. However, across the Congo Basin
this data is rare, or difficult to access. Despite the difficulties in
assessing hidden disturbances caused by deforestation, poorly recorded
official assessments or simply those invisible to standard remote
sensing techniques \citep{peres2006}, remote sensing generally remains
the best way to map and quantify LULCC \citep{houghton2012}. Despite
these challenges in recovery and validation we present a first analysis
of LULCC using historical aerial survey data for the central Congo
Basin. We used a combination of Structure-from-Motion and a
convolutional neural net deep learning approach to map deforestation
across a large extent, providing a long term view of LULCC in the
central Congo Basin

\hypertarget{data-recovery-challenges}{%
\subsection{Data recovery challenges}\label{data-recovery-challenges}}

In our study the archive data recovered was limited to contact prints
and therefore did not represent the true resolution of the original
negative. In addition, analogue photography clearly produces a distinct
softness compared to digital imagery (Figure 4). Despite favourable
nadir image acquistions \citep{verhoeven2017} image softness combined
with illumination effects between flight paths, and the self-similar
nature of vast canopy expanses \citep{park2019, simini2010, sole1995},
limited our ability to provide wall-to-wall coverage of the entire
dataset containing 334 images. Few man made features in the scenes also
made georeferencing challenging. Although the village of Yangambi
provided a range of buildings as (hard-edge) references, other areas
within the central Congo Basin might have fewer permanent structures and
would require the use of soft-edged landscape features (e.g.~trees,
river outflows). Research has shown that soft-edged features can help
georeference scenes even when containing few man-made features
\citep{hughes2006}, however accuracy becomes compromised. Our two step
georeferencing approach resulted in a referencing accuracy of
\textasciitilde{}4.7 \(\pm\) 4.3 m across reference points. However, it
shoud be noted that referencing accuracy of the final scene is less
constrained toward the edges of the scene.

\hypertarget{land-use-land-cover-change-above-ground-carbon}{%
\subsection{Land-use, Land-Cover change \& Above Ground
Carbon}\label{land-use-land-cover-change-above-ground-carbon}}

When classifying our orthomosaic into forest and non-forest states we
favoured a deep learning supervised classification using a Convoluted
Neural Network over manual segmentation to guarantee an
``apples-to-apples'' comparison between the historical and the
contemporary GFC forest cover maps. We acknowledge that both
classifications use different features, i.e.~spectral or spatial data,
but attain a similarly high accuracy of 97\% and 99\%
\citep{hansen2013}, for the GFC and our CNN map respectively. Despite a
number of challenges we were able to map and quantify a vast continuous
area of tropical forest, with limited manual annotation requirements.

Our analysis shows that the majority of deforestation around Yangambi
happened toward the late 1950's (162 km\textsuperscript{2}).
Considerable regrowth has occured since the aerial survey was executed
(99 km\textsuperscript{2}), and socio-economic instability prevented
further large scale forest exploitation. In particular, many plantations
have reached maturity and forest has re-established in previously
cleared or disturbed areas. The majority of this forest recovery takes
the form of isolated patches of forest but is offset by further
deforestation of previously untouched forest. Generally, the function
and structure of forests can be influenced by forest edges that are
located up to 1km away however most effects are pronounced within the
first 300m from the edge \citep{gascon2000}. Our analysis of edge
effects on AGC has shown that the influence is negligible 200m away from
the edge. \citet{phillips2006} have shown similar weak responses to edge
effects in the Amazon forest. Due to a lack of data on the extent
(depth) of edge effects and their influence on AGC beyond 200m we did
not include any estimates of carbon loss or gain within these zones.
However, it must be stated that the influence of edges throughout the
landscape was not marginal. as these areas would account for 132
km\textsuperscript{2}. Thus edges could have a substantial negative
\citep{brinck2017} or positive \citep{reinmann2017} influence on AGC.
Similarly, uncertainties in how to explicitly correct for plantations in
the landscape present a further challenge. Thus although our estimates
are only indicative they do underscore the important influence of
landscape structure when carbon accounting. However, our findings do not
indicate that deforestation in Congo basin is declining, on the
contrary.

Over the past half century there has been a clear shift in land use in
Yangambi (Figure 5). Land use has shifted away from for example a
regular fishbone deforestation pattern that emerges when (large scale)
agricultural interests dominate the landscape \citep{arima2015}, to a
more fragmented landscape (Figure 5D). The latter former? is consistent
with historical land management at the time of the aerial survey
\citep{bustillo2018}. These regular patterns have since been reversed
because of a decrease in large scale intensive agriculture, replaced by
ad-hoc small scale subsistence farming with large perimeter to area
relationships (i.e.~ragged edges). Consequently, edge effects in the
current landscape are far more pronounced than in the historical scene.

Visual inspection of the images also suggests that reforestation within
the historically cleared areas and experimental plots is not necessarily
limited to areas far removed from more densely populated areas. For
example, large reforested areas exist close to the Congo stream and
Yangambi village itself (Fig. 5). Here, regional political components,
such as land leases and large scale ownership could have played a role
in safeguarding some of these areas for rewilding or sustainable
management \citep{arima2014, larson2011}. Despite widespread
anthropogenic influences throughout the tropics \citep{lewis2015} the
retention of these forested areas show the potential of explicit or
implicit protective policy measures (e.g.~INERA concessions,
\citet{bustillo2018}) on a multi-decadal time scale. Forest regrowth in
non-continuous areas within Yangambi could increase landscape
connectivity and help increase biodiversity \citep{vandeperre2018}.

Our analysis therefore provides an opportunity to highlight and study
those regions that have previously suffered confirmed long-term
disturbances, and those that have been restored since. Assessments of
old plantations and recovering clear-cut forests can serve as a guide to
help estimate carbon storage capacity and forest recovery rates in
managed and unmanaged conditions
\citep{gourlet-fleury2013, sader1988, achard2014} over the mid- to
long-term, in support of rewilding and general forest restoration
\citep{arima2014, larson2011, vandeperre2018}. In addition, mapping
long-term edge effects can further support research into issues such as
receding forest edges \citep{gascon2000}.

\hypertarget{canopy-structure-foto-texture-analysis-2}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-2}}

Finally the FOTO technique used to quantify relationships between canopy
structure and forest characteristics rendered no valuable insights of
either the historical orthomosaic or recent Geo-Eye scene. Similarly
weak correlations were found previously by \citet{solorzano2018}. In
contrast site based texture metric statistics did show correspondence
between historical and contemporary satellite imagery. However, none of
them were either consistent or significant. Although visual
interpretation shows distinctly different canopy structures (Figure 5)
the differences in how resolution is defined and issues related to image
quality prevented us from quantifying these further. Unlike large scale
studies by \citet{Ploton2012} we could not scale this technique to
historical data. We advise that future valorisation efforts should
preferentially work from foto negatives (if available) to ensure optimal
data quality in resolution, contrast and overall sharpness.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Given the impact of tropical forest disturbances on atmospheric CO\_2
emissions, biodiversity and ecosystem productivity accurate long term
reporting on land-cover and land-use change especially in the
pre-satellite era is an imperative. Our analysis of historical aerial
survey images of the Central Congo Basin provides a window into the
state of the forest at the start of the anthropocene. Efforts to
quantify canopy texture and their link to carbon storage had limited to
no success.We have shown that historical aerial survey data can be used
to quantify land-use and land-cover change using a highly automated
workflow to quantify the structure and extent of forest cover change
that can help assess the impact of fragmentation on above ground carbon
stocks. We also show a shift from previously highly structured
industrial deforestation of large areas, to discrete smallholder
clearing for farming, increasing landscape fragmentation but also
opportuntiy for substantial regrowth. Efforts to quantify canopy texture
and their link to carbon storage had limited to no success. As such, our
analysis provides insights into the state of rarely studied tropical
forests and the rate at which deforestation and reforestation has taken
place over a multi-decadal scale in the central Congo basin providing a
useful historical context to land-cover and land-use change spatial
patterns for interpreting past and ongoing field research in the area.

\hypertarget{additional-information-and-declarations}{%
\section{Additional Information and
Declarations}\label{additional-information-and-declarations}}

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

Hufkens et al.~(2019). A curated dataset of aerial survey images over
the central Congo Basin, 1958. Zenodo: doi.org/10.5281/zenodo.3547767.
All data not included in the latter repository can be found bundled with
the analysis code as listed below. Proprietary datasets (i.e.~Geo-Eye
data) are not shared, but purchase order numbers allow for acquisition
of these datasets to ensure reproducibility. The \citet{hansen2013}
image data is freely available from
\href{https://earthenginepartners.appspot.com/science-2013-global-forest/download_v1.6.html}{the
GFC project website}.

\hypertarget{code-availability}{%
\subsection{Code availability}\label{code-availability}}

All analysis code is available as an R \citep{rcoreteam2019} project
(\url{https://github.com/khufkens/orthodrc}). The analysis relied
heavily on the `raster' \citep{hijmans2019}, `RStoolbox'
\citep{leutner2019}, and `landscapemetrics' \citep{hesselbarth2019}
packages, while post-processing and plotting was facilitated by the
`tidyverse' ecosystem \citep{wickham2017}, `ggthemes'
\citep{arnold2019}, `scales' \citep{wickham2018} and `cowplot'
\citep{wilke2019}. Additional plotting elements were formatted or
provided by `sf' \citep{pebesma2018} and `rnaturalearth'
\citep{south2017} packages, respectively. I'm grateful for the
contributions to the scientific community by the developers of these
packages.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
%
% % Only for the journal Methods and Protocols:
% % If you wish to submit a video article, please do so with any other supplementary material.
% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This research was supported through the Belgian Science Policy office
COBECORE project (BELSPO; grant BR/175/A3/COBECORE). KH acknowledges
funding from the European Union Marie Skłodowska-Curie Action (project
number 797668). I'm grateful for the support of Thales D'Haulleville
volunteering his time in scanning the images.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{K.H. conceived and designed the study, analyzed the data, prepared
figures and tables, authored, reviewed and approved the final draft of
the manuscript. T.d.H. scanned all image data. E.K. and T.d.H. provided
plot based AGC estimates. T.D., K.J., E.K, H.B., P.S., F.V.,S.M., M.A.,
J.V.D.B., H.V and L.W. reviewed the final manuscript.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest. The founding sponsors had
no role in the design of the study; in the collection, analyses, or
interpretation of data; in the writing of the manuscript, an in the
decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
% \reftitle{References}
% \begin{thebibliography}{999}
% % Reference 1
% \bibitem[Author1(year)]{ref-journal}
% Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149, DOI.
% % Reference 2
% \bibitem[Author2(year)]{ref-book}
% Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58, ISBN.
% \end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{bibliography.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}