### 2020-01-24 manuscript/main.tex

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[remote sensing,article,submit,moreauthors,pdftex]{mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

\usepackage{longtable}

% see https://stackoverflow.com/a/47122900

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Historical aerial surveys map long-term changes of forest cover and
structure in the central Congo Basin}

% Authors, for the paper (add full first names)
\Author{Koen Hufkens$^{1,2,*}$\href{https://orcid.org/0000-0002-5070-8109}{\orcidicon}, Thalès de Haulleville$^{3}$, Elizabeth Kearsley$^{1}$, Kim Jacobsen$^{1,3}$, Hans Beeckman$^{3}$, Piet Stoffelen$^{4}$, Filip Vandelook$^{4}$, Sofie Meeus$^{4}$, Michael Amara$^{5}$, Leen Van Hirtum$^{5}$, Jan Van den Bulcke$^{1}$, Hans Verbeeck$^{1}$, Lisa Wingate$^{2}$}

% Authors, for metadata in PDF
\AuthorNames{Koen Hufkens, Thalès de Haulleville, Elizabeth Kearsley, Kim Jacobsen, Hans Beeckman, Piet Stoffelen, Filip Vandelook, Sofie Meeus, Michael Amara, Leen Van Hirtum, Jan Van den Bulcke, Hans Verbeeck, Lisa Wingate}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Ghent University, Ghent, Belgium; \\
$^{2}$ \quad INRA, UMR ISPA, Villenave d'Ornon, France; \\
$^{3}$ \quad Royal Museum for Central Africa, Tervuren, Belgium; \\
$^{4}$ \quad Botanic Garden Meise, Meise, Belgium; \\
$^{5}$ \quad National Archives of Belgium, Brussels, Belgium; \\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:koen.hufkens@gmail.com}{\nolinkurl{koen.hufkens@gmail.com}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Given the impact of tropical forest disturbances on atmospheric carbon
emissions, biodiversity and ecosystem productivity accurate long term
reporting of Land-Use and Land-Cover (LULC) change in the pre-satellite
era (\textless{} 1972) is an imperative. Here, we use a combination of
historical (1958) aerial photography and contemporary remote sensing
data to map long-term changes in the extent and structure of the
tropical forest surrounding Yangambi (DR Congo), in the central Congo
Basin. Our study leveraged Structure-from-Motion and a novel
Convolutional Neural Network based LULC classifier, using synthetic
landscapes based image augmentation, to map historical forest cover
across a large sub meter resolution orthomosaic (\textasciitilde{}82800
ha) geo-referenced to \textasciitilde{}4.7 \(\pm\) 4.3 m. A comparison
with contemporary LULC data showed a shift from previously highly
regular industrial deforestation of large areas, to discrete smallholder
farming clearing, increasing landscape fragmentation but also providing
opportunties for substantial forest regrowth. Efforts to quantify long
term canopy texture changes and a link to above ground carbon had
limited to no success. Our analysis provides methods and insights into
key spatial and temporal patterns of deforestation and reforestation at
a multi-decadal scale, providing a historical context for past and
ongoing forest research in the area.}

% Keywords
\keyword{Aerial survey, data recovery, CNN, Deep Learning, SfM, Congo Basin}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{wrapfig} \usepackage{graphics} \usepackage{setspace} \doublespacing \usepackage{gensymb} \usepackage{booktabs}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Tropical ecosystem services are severely impacted by deforestation and
forest degradation \citep{hansen2013, houghton2012, tyukavina2015}. Not
only does tropical forest Land-Use and Land-Cover Change (LULCC)
constitute 10 to 15\% of the total global carbon emissions
\citep{vanderwerf2009}, changes in forest fragmentation affect the
forest structure and function
\citep{fauset2017, brinck2017, didham1999}. Strong fragmentation effects
decreases the number of large trees along forest edges
\citep{laurance2000, magnago2015}, while species composition and
biodiversity are equally negatively affected
\citep{poorter2006, barlow2016, vandeperre2018}. Estimates show that
31\% of carbon emissions are caused by edge effects alone
\citep{brinck2017}.

Accurate estimates of LULCC and forest canopy structure are therefore
imperative to estimate carbon emissions and other ecosystem services
\citep{houghton2012, hansen2013}. Remote sensing products have been key
inputs in LULCC assessments as they provide accurate spatial information
to help estimate carbon emissions \citep{mitchard2018, hansen2013}. More
so, high resolution aerial images provide scientists tools to monitor
forest extent, structure and carbon emissions as canopy texture is
linked to above ground biomass
\citep{Ploton2012, Couteron2005, Barbier2010}. Yet, most of these
estimates are limited in time to recent decades
\citep{defries2002, houghton2012, hansen2013, achard2014}.

Historical estimates \textbf{of Land-Use and Land-Cover (LULC)}, in the
pre-satellite era (\textless{} 1972), exist but generally rely on
non-spatially explicit data (i.e socio-economic data)
\citep{houghton2012, ramankutty1999, defries2002, sader1988}. Efforts
have been made to use other geo-spatial data sources such as historical
maps \citep{willcock2016}, declassified CORONA satellite surveillance
data across the US and central Brazil \citep{song2015} as well as aerial
surveys in post World War II German \citep{nita2018}. Survey data across
the African continent is less common, inaccessible or both. Some studies
do exist, as \citet{buitenwerf2012} and \citet{hudak1998} used aerial
survey images to map vegetation changes in South African savannas,
whilst \citet{frankl2015} and \citet{nyssen2016} mapped the Ethiopian
highlands of 1930's.

Across the central Congo Basin most of these historical images were
collected within the context of national cartographic efforts by the
``Institut Géographique du Congo Belge'' in Kinshasa (then
Léopoldville), DR Congo. Despite the existence of large archives of
aerial survey imagery (Figure 1, Appendix Figure 3), as of yet, no
studies have valorized these data. The lack of a consistent valorization
effort is unfortunate as the African rainforest is the second largest on
Earth and covers \textasciitilde{}630 million ha, representing up to 66
Pg of carbon storage \citep{Lewis2009b}, and currently loses forest at
an increasing pace \citep{butsic2015}. Given the impact of LULCC on the
structure and functioning of central African tropical forests, and their
influence on both carbon dynamics \citep{Kearsley2013} and biodiversity
\citep{vandeperre2018}, accurate long term reporting of historical
forest cover warrants more attention \citep{willcock2016}.

Here, we use a combination of historical aerial photography and
contemporary remote sensing data to map long-term changes in the extent
and structure of the tropical forest surrounding Yangambi (DR Congo), in
the central Congo Basin, effectively linking the start of the
antrophocene \citep{lewis2015} with current assessments. Yangambi was,
and remains, a focal center of forest and agricultural research and
development in the central Congo Basin. Past research in the region
allows for thorough assessment of land-use and land-cover change from a
multi-disciplinary point of view, confronting us with complex
deforestation and land-use patterns.

We leverage Structure-from-Motion to generate a large orthomosaic of
historical imagery and develop a Convolutional Neural Network based
forest cover mapping approach, based upon a semi-supervised generated
dataset extensively leveraging data augmentation. Our methodology aims
to provides a historical insight into important LULCC spatial patterns
in Yangambi, such as fragmentation and edge complexity. We further
contextualize the influence of changes in the forest's life history on
past and current research into Above Ground Carbon (AGC) storage
\citep{Kearsley2013} and biodiversity \citep{vandeperre2018} in the
central Congo Basin. Our fast scalable mapping approach for historical
aerial survey data, using limited supervised input, would further
support long-term land-use and land-cover change analysis across the
central Congo Basin.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{historical-data-acquisition}{%
\subsection{Historical data
acquisition}\label{historical-data-acquisition}}

Data for the central Congo Basin region, surrounding Kisangani, were
collected in several flights during the dry season of 1958 and 1959
(from 8/01/1958 to 20/02/1958 and from 28/12/1958 to 9/01/1959
respectively, see Appendix Figure 1) to generate topographic maps of the
area, supervised by the ``Institut Géographique du Congo Belge'' in
Kinshasa, DR Congo (then Léopoldville). Black-and-White infrared images
(0.4 - 0.9 \(\mu\)m) were gathered along flight paths running mostly
from west to east, between \textbf{09:00 and 11:00} local time. Along a
flight path continuous images were taken using a Wild Heerburgg RC5a
(currently Leica Geosystems) with an Aviogon lens assembly (114.83mm / f
5.6, with a 90\degree view angle) resulting in square photo negatives of
180 by 180 mm. Flights were flown at an average absolute altitude of
\textasciitilde{}5200 m above sea level, covering roughly 18 530
km\textsuperscript{2} at an approximate scale of 1/40 000. The use of
the integrated autograph system ensured timely acquisition of pictures
with a precise overlap (\textasciitilde{}1/3) between images. This large
overlap between images together with flight parameters would allow
post-processing, using stereographs, to create accurate topographic
maps. Original data from this campaign are stored in the
\href{https://www.africamuseum.be/en}{Royal Museum for Central Africa}
in Tervuren, Belgium (Figure 1).

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{./figures/print_examples}

}

\caption{A box of historical aerial photographs (left) and a single aerial photograph (right) showing part of the Congo river. Note the meta-data provided in the right side margin of the image, such as acquisition time and fleight height.}\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{site-selection}{%
\subsection{Site selection}\label{site-selection}}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{./figures/flight_paths}

}

\caption{Overview of the historical flight paths during aerial photo acquisition and ancillary data used in this study. The bounding box of the orthomosaic data presented in this study is shown as a rectangle (23x36 km). The outline of a recent high-resolution Geo-eye panchromatic image is shown as a dashed dark grey rectangle (10x10km). The location of various permanent sampling plots are shown as x, +, and open squares and triangles for the mixed, mono-dominant and edge plots respectively. The grey polygon delineates the current day Yangambi Man and Biosphere reserve. The inset, top right, situates the greater Yangabmi region (white rectangle) with the DR Congo. The full flight plan and details are shown in Appendix Figure 1 and 2}\label{fig:unnamed-chunk-2}
\end{figure}

We prioritised flight paths and images that contained current day
permanent sampling plots, larger protected areas, and past agricultural
and forest research facilities (Figure 2). This selection provides a
comprehensive mapping of the Yangambi area and the life history of the
forest surrounding it. Thereafter, we selected flight paths 1 through 11
for digitization. From this larger dataset of 334 images we selected 74
survey images for orthomosaic compositing and further analysis. All the
selected images stem from the flight campaign made during January and
February of 1958. The area includes the Yangambi village, 20
contemporary permanent sampling plots \citep{Kearsley2013}, past and
present agricultural experimental plots \citep{bauters2015} and large
sections of the Yangambi
\href{http://www.unesco.org/new/en/natural-sciences/environment/ecological-sciences/biosphere-reserves/africa/democratic-republic-of-the-congo/yangambi/}{UNESCO
Man and Biosphere reserve} surrounding to the west and east of the
village. Although not formally mosaicked we provided a full dataset of
pre-processed images using the cropping and normalization routines
described below. The latter data was not used in subsequent LULCC
analysis, but has been archived and made available to the public
separately (see code \& data availability statement below).

\hypertarget{digitization-and-data-processing}{%
\subsection{Digitization and data
processing}\label{digitization-and-data-processing}}

All selected images, covering the Yangambi area, were contact prints as
original negatives of the prints were not available. Images were scanned
at a resolution exceeding their original resolution (or grain) at the
maximal physical resolution of an Epson A3 flatbed scanner (i.e.~2400
dpi or 160 MP per image) and saved as lossless tiff images. Data were
normalized using contrast limited histogram equalization
\citep{zuiderveld1994} with a window size of 32 and a clip limit of 1.5.
Fiduciary marks were used to rectify and downsample the images into
square 7700x7700 pixel images (\textasciitilde{}1200 dpi, 81 MP). This
resulted in a dataset with digital images at a resolution that remained
above the visible grain of the photographs, whilst the reduced image
size facilitated easier file handling and processing speed.

Data was processed into a georeferenced orthomosaic using a
Structure-from-motion (SfM, \citet{ullman1979}) approach implemented in
\href{http://www.agisoft.com}{Agisoft Metashape} version 1.5.2 (Agisoft
LLC, St.~Petersburg, Russia). An orthomosaic corrects remote sensing
data to represent a perfectly downward looking image, free from
perspective distortions due to topography and camera tilt. Using the SfM
technique features, areas in images with a large degree of similarity,
are matched across various images to reconstruct a three dimensional
scene (topography) from two-dimensional image sequences. During the SfM
analysis we masked clouds, glare or large water bodies such as the Congo
river.

We calculated the orthomosaic using a low resolution point cloud and
digital elevation map (DEM). Additional ground control points were
provided to assist in the referencing of image and constrain the
optimization routine used in the SfM algorithm. Ground control points
consisted of permanent structures which could be verified in both old
and new aerial imagery (i.e.~ESRI World Imagery) \textbf{and consisted
of corner points of build structures (e.g.~a building, bridge or
swimming pool etc.).} Although clouds were removed during the SfM
routine we did not mask all clouds in the final orthomosaic to maximize
forest coverage. The final scene was cropped, to provide consistent wall
to wall coverage of the reconstructed scene. The orthomosaic was
exported as a geotiff for further georeferencing in QGIS
\citep{qgisdevelopmentteam2018} using the georeferencer plugin (version
3.1.9) and additional ESRI World Imagery high resolution reference data.
We used 3\textsuperscript{th} degree polynomial and 16 ground control
points to correct the final image. Ground control points, raw image data
and final processed image are provided in addition to measures of
uncertainty such as mean and median error across all ground control
points. All subsequent analysis are executed on the final geo-referenced
orthomosaic or subsets of it.

\hypertarget{land-use-and-land-cover-change}{%
\subsection{Land-Use and Land-Cover
Change}\label{land-use-and-land-cover-change}}

\hypertarget{classifying-land-use-and-land-cover}{%
\subsubsection{Classifying Land-Use and
Land-Cover}\label{classifying-land-use-and-land-cover}}

\hypertarget{model-training}{%
\paragraph{\texorpdfstring{\emph{Model
training}}{Model training}}\label{model-training}}

We automatically delineated all natural forest in the historical data,
thus excluding tree plantations, thinned or deteriorated forest stands
which \textbf{showed visible canopy cover loss}, fields and buildings.
We used the Unet Convolutional Neural Net (CNN, \citet{ronneberger2015})
architecture implemented in Keras \citep{chollet2015} with an
efficientnetb3 pre-processing backbone \citep{yakubovskiy2019} running
on TensorFlow \citep{martinabadi2015} to train a binary classifier
(i.e.~forest or non-forested). Training data were collected from the
orthomosaic by randomly selecting 513 pixel square tiles from
\textbf{locations within} homogeneous forested or non-forested polygons
in the historical orthomosaic \textbf{(Figure 5). Separate polygons were
selected for training, testing and validation purposes. Validation
polygons were sampled 300 times, while both testing and validation
polygons were sampled at 100 random locations. Tiles extracted from
locations close to the polygon border at times contained mixed cover
types. Tiles with mixed cover types were removed from the list of source
tiles (Table 1).} Homogeneous \textbf{source} tiles were combined in
synthetic landscapes using a random gaussian field based binary mask
(Figure 3). We generated 5000 synthetic landscapes for training, while
500 landscapes were generated for both the validation and the testing
datasets for a total of 6000 synthetic landscapes. In order to limit
stitch line misclassifications, along the seams of mosaicked images, we
created synthetic landscapes with different forest tiles to mimick
forest texture transitions. We applied this technique to 10\% of the
generated synthetic landscapes \textbf{(across training, validation and
testing data)}.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Number of source tiles used for the generation of synthetic landscapes.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  & training & testing & validation\\
\midrule
forest & 300 & 100 & 100\\
non-forest & 294 & 92 & 84\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/synthetic_landscape}

}

\caption{An example synthetic landscape, combining homogeneous forest and non-forest images into a patchy landscape using random gaussian field based masks. The left panel shows a combined synthetic landscape, while the right panel shows the corresponding forest (black) and non-forest (white) labels.}\label{fig:unnamed-chunk-4}
\end{figure}

The CNN model was trained for 100 epochs with a batch size of 30 using
Adam optimization \citep{kingma2017}, maximizing the
Intersect-over-Union (IoU) using Sørensen--Dice \citep{dice1945} and
categorical cross-entropy loss functions. Data augmentation included
random cropping to 320 pixel squares, random orientation, scaling,
perspective, contrast and brightness shifts and image blurring. The
optimized model was used to classify the complete orthomosaic using a
moving window approach with a step size of 110 pixels and a majority
vote \textbf{(\textgreater{} 50\% agreement)} across overlapping areas
to limit segmentation edge effects. \textbf{In addition, we provide raw
pixel level classification agreement data for quality control purposes
(see Data availability below).} We refer to Figure 4 for a synoptic
overview of the full deep learning workflow.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/cnn_diagram}

}

\caption{A diagram of the deep learning workflow followed in training a binary forest / non-forest cover convolutional neural net U-Net model to generate our forest cover map.}\label{fig:unnamed-chunk-5}
\end{figure}

\hypertarget{model-validation}{%
\paragraph{\texorpdfstring{\emph{Model
validation}}{Model validation}}\label{model-validation}}

\textbf{We report the CNN accuracy based upon the IoU of our
out-of-sample validation dataset of synthetic landscapes. In addition,
we report confusion matrices for all pixels across the homogeneous
validation polygons, as well as the training and testing polygons (see
Figure 5). Furthermore, we used the first acquisition of a recent
pan-chromatic Geo-Eye 1 stereo pair (Geo-Eye, Thornton, Colorado,
U.S.A., order 737537, 2011-11-11 08:55 GMT or 09:55 local time) to
classify and assess the robustness of the CNN algorithm on contemporary
remote sensing imagery with similar spectral and spatial
characteristics. We used the Global Forest Change version 1.6 (GFC, tile
10N-020E) \citep{hansen2013} map data, capturing forest loss up to 2011,
to quantify accuracy on downsampled CNN Geo-Eye classification results.
Once more, we report the confusion matrix of between the GFC and CNN
derived forest cover maps, masking clouds and cloud shadows. To
summarize confusion matrices we report accuracy as:}

\(Accuracy = \frac{(TP + TN)}{(TP + TN + FP + FN)}\)

\textbf{in which TP, TN, FP, FN are True Positive, True Negative, False
Positive and False Negative, respectivelly.}

\hypertarget{characterizing-long-term-change}{%
\subsubsection{Characterizing long term
change}\label{characterizing-long-term-change}}

To map long term LULCC in the Yangambi region we used the contemporary
Global Forest Change version 1.6 (GFC, tile 10N-020E) map data
\citep{hansen2013}. Using the \textbf{GFC} data we calculated the latest
state of the forest with respect to the conditions at the start of 1958,
60 years earlier. In our analysis we only included \textbf{GFC} pixels
which recorded no \textbf{forest} loss throughout the whole 2000 - 2018
period. As the resolution of the historical forest classification
exceeds that of the GFC map we downsampled our historical forest cover
data to 30 m GFC resolution \textbf{using a nearest neighbour approach}.
We masked out all water bodies using the Global Forest Change survey
data mask layer, and limited the analysis to the right bank of the Congo
river. We provide summary statistics of historical and contemporary
forest loss and regrowth. \textbf{We map persistent forest loss after
1958, regrowth after loss in 1958, recent forest loss and long term
(stable) forest cover.}

\hypertarget{landscape-fragmentation-above-ground-carbon-estimates}{%
\subsubsection{Landscape fragmentation \& Above Ground Carbon
estimates}\label{landscape-fragmentation-above-ground-carbon-estimates}}

To quantify changes in the structure of forest cover and its
disturbances we used spatial pattern analysis landscape fragmentation
metrics \citep{dale1999}. We report the ratio of edge to area and the
fractal index to quantify landscape complexity of forest disturbances.
Statistics were calculated for all forest disturbance patches larger
than 1 ha and smaller than the 95\textsuperscript{th} percentile of the
patch size distribution using the R package landscapemetrics
\citep{hesselbarth2019}. We provide mean and standard deviation on edge,
area, their ratio and fractal dimension for both the historical and
contemporary \citet{hansen2013} forest cover maps.

We estimated above ground carbon (AGC) losses and gains over time using
plot based averages of recent inventory data at permanent sampling plots
in the area (Figure 2). We refer to \citet{Kearsley2013} for the survey
method and allometric relations used to scale the survey data. Unlike
standard square 1 ha plots, edge plots were set back 200 m from forest
edges and were 50x200 m, with the 50 m side of the plot along the forest
edge and continuing 200 m into the forest (Appendix Table 1). We further
confirmed that forest edge plots did not show a significantly different
AGC compared to those of non-edged / mixed forest plots (Mann Whitney U
test, p \textless{} 0.05). Thus it was not necessary to explicitly
quantify changes in AGC caused by edge effects. Moreover, we used the
mean values of the mixed forest as representative for potential AGC
losses. Despite the challenges inherent in quantifying AGC for forest
edges we mapped the total extent of the edges in the contemporary
landscape. To align our landscape analysis with exploratory analysis of
the survey data we used a buffer of 200 m to estimate the extent of
forest edges and patches, up to the location of forest edge plots.

Surveys of old plantations show a large variation in AGC, depending on
age and the crop type. For example, the AGC values varied from 168.67 to
86.55 Mg C ha\textsuperscript{-1}, for \emph{Hevea brasiliensis} (rubber
tree) and \emph{Elaeis guineensis} (oil palm) plots respectively
(\citet{bustillo2018}, personal communications). These higher values are
in line with the mixed AGC estimates in the area, while the palm
plantations resemble old-regrowth values (81.8 Mg C
ha\textsuperscript{-1}, see \textbf{Appendix Table 1}). We therefore use
both the estimates of old-regrowth and mixed forest to estimate AGC for
regrowth. We did not have sufficient data to account for individual
changes in AGC across plantations.

\hypertarget{canopy-structure-foto-texture-analysis}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis}}

We compared the structure of the canopy both visually and using Fourier
Transform Textural Ordination (FOTO, \citet{couteron2002}). FOTO uses a
principal component analysis (PCA) on radially averaged 2D Fourier
spectra to characterize canopy (image) texture. The FOTO technique was
first described by \citet{couteron2002} to quantify canopy stucture in
relation to biomass and biodiversity, and can be used across multiple
scenes using normalization \citep{Barbier2010}.

We used the standard FOTO methodology with fixed zones, instead of the
moving window approach. The window size was set to the same size (187
pixels or \textasciitilde{}150 m) as used in the moving window analysis
above. To account for illumination differences between the two scenes we
applied histogram matching. No global normalization was applied, as the
scene was processed as a whole. PC values from this analysis for all
permanent sampling plots in both image scences were extracted using a
buffer with a radius of 50 m.

For both site based and scene analysis we correlated PC values with
permanent sample plot inventory data such as stem density, above ground
biomass and tree species richness. Additional comparisons are made
between contemporary Geo-Eye data and the historical orthomosaic derived
PC values. Due to the few available permanent sampling plots in both
scenes we used a non-parametric paired signed rank (Wilcoxon) test
\citep{wilcoxon1945} to determine differences between the PC values of
the Geo-Eye and historical orthomosaic image scenes across mono-dominant
and mixed forest types. In all analysis, mono-dominant site 4 was
removed from the analysis due to cloud contamination.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{orthomosaic-construction}{%
\subsection{Orthomosaic construction}\label{orthomosaic-construction}}

Our analysis provides a first spatially explicity historical composite
of aerial survey images \textbf{in support of mapping land-use and
land-cover} within the Congo Basin. The use of high resolution
historical images combined with SfM image processing techniques allowed
us to mosaic old imagery across a large extent. The final orthomosaic
composition of the Yangambi region \textbf{provided} an image scene
covering approximately 733 million pixels across
\textbf{\textasciitilde{}82800 ha} (\textasciitilde{}23x36 km, Figure
2). The overall accuracy of the SfM orthomosaic composition was 0.88
m/pixel using the sparse cloud DEM for corrections at 45.8 m/pixel. The
resulting georeferenced scene had a spatial accuracy of approximately
\textbf{23 m}. Further georeferencing outside the SfM workflow reduced
the mean error at the ground control points to 5.3 \(\pm\) 4.9 px
(\textasciitilde{}4.7 \(\pm\) 4.3 m), with a median error of 2.9 px (2.6
m). The orthomosaic served as input for all subsequent LULCC analysis
with all derived maps provided with the manuscript repository (see data
\& code availability statements below).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/orthomosaic_maps}

}

\caption{Overview of the final orthomosaic of the greater Yangambi region (A) as well as the forest cover classification uncertainty (B) used to generate the final Land Land Cover map (see Figure 6). Homogeneous polygons used in the generation of the synthetic landscape for Convolutional Neural Network training, testing and validation are marked as dashed and full lines for forest and non-forest regions, respectively. Training, testing and validation regions are denoted in panel B in green, red and orange, respectivelly. Black polygon outlines denote cloud and image stitch line regions which were manually excluded from analysis, but retained in calculation of validation statistics (see Table 2).}\label{fig:unnamed-chunk-6}
\end{figure}

\hypertarget{land-use-and-land-cover-classification}{%
\subsection{Land-use and Land-Cover
Classification}\label{land-use-and-land-cover-classification}}

\hypertarget{cnn-model-validation}{%
\subsubsection{CNN model validation}\label{cnn-model-validation}}

\textbf{The CNN deep learning classifier reached an
Intersection-over-Union accuracy of 97\% on the detection of disturbed
forest in the out-of-sample (validation) synthetic landscape data. Using
all pixels within the validatoin polygons (Figure 5) showed a similar
accuracy value of \textasciitilde{}98\%. Using all polygons across the
scene, including those used in the generation of testing and training
synthetic landscapes, increased the accuracy to \textasciitilde{}99\%
(Table 2). A comparison with recent pan-chromatic Geo-Eye data shows
good agreement, with an accuracy of \textasciitilde{}87\% across all
pixels, between the landsat based GFC data and downscaled CNN results
(Table 2 and Figure 7).}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-7}Confusion matrix showing \% agreement between forest / non-forest classes using a Convolutional Neural Network (CNN) across previously selected homogenous areas.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
\multicolumn{1}{c}{CNN} & \multicolumn{2}{c}{Geo-Eye (full scene)} & \multicolumn{2}{c}{Aerial Survey (all polygons)} & \multicolumn{2}{c}{Aerial Survey (validation polygons)} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
  & non-forest & forest & non-forest & forest & non-forest & forest\\
\midrule
non-forest & 10.17 & 8.55 & 97.4 & 0.19 & 98.59 & 0.0\\
forest & 3.75 & 77.52 & 2.6 & 99.81 & 1.41 & 100.0\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Accuracy}}\\
\hspace{1em} &  & 87.70 &  & 98.61 &  & 99.3\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/forest_cover_map}

}

\caption{Overview of the final Land Use Land Cover Change map, a detailed inset of both the underlying orthomosaic (B) and the derived land use land cover change map displayed as the difference between the Convolutional Neural Network based classification orthomosaic and the recent Landsat based forest cover map by Hansen et al. 2013 (C).}\label{fig:unnamed-chunk-8}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/visual_comparison_classifiers}

}

\caption{Convolutional Neural Network (CNN) based forest cover classification results (B) as run on a recent (2011) Geo-Eye panchromatic image (A). We show the difference between the Convolutional Neural Network based classification and a recent Landsat based forest cover map by Hansen et al. 2013. Full white outlines denote cloud contamination, the dashed rectangle shows a location where the CNN outperforms the Landsat based forest classification.}\label{fig:unnamed-chunk-9}
\end{figure}

\hypertarget{long-term-changes-in-lulc-and-above-ground-carbon}{%
\subsection{Long term changes in LULC and Above Ground
Carbon}\label{long-term-changes-in-lulc-and-above-ground-carbon}}

Scaling our classifier to the whole historical orthomosaic we detected
\textbf{\textasciitilde{}16200 ha} (or \textasciitilde{}20\% of the
scene) of disturbed forests. A large fraction of the disturbed area was
restored in the period between the two \textbf{acquisition periods}. In
total, \textbf{9918 ha}, or little over half of the affected forest was
restored (Figure 5C-D, dark blue). Recent deforested areas, as
registered through satellite remote sensing, approximate \textbf{8776 ha
(Table 3, Figure 6 - light green)}.

Recent deforestation follows a distinctly different pattern compared to
historical patterns. Historical deforestation showed a classical
fishbone pattern for forest clearing with very sharp edges, while
current patterns are patchy and ad-hoc (\textbf{figure 6C}, dark blue
and green colours respectively). These differences are reflected in the
analysis of landscape metrics of forest loss. Between the historical and
contemporary LULCC maps we see an increase in small disturbances, as
indicated by the decreasing area of the mean patch size, down to
\textasciitilde{}1.86 \(\pm\) 0.75 ha from \textasciitilde{}5.25 \(\pm\)
5.02 ha historically. Perimeter lengths were longer historically, at
1451 \(\pm\) 943 m, compared to contemporary landscapes
\textasciitilde{}921 \(\pm\) 362 m (Table 3). This shift in perimeter
area ratio led to a similar change in the fractal index, slightly
increasing in value to 1.1 \(\pm\) 0.05 from 1.09 \(\pm\) 0.04 over
time. Values closer to a fractal index of 2 suggest a more complex
(fragmented) landscape.

A comparison of forest edge plots with mixed forest plots showed no
signficant difference in AGC, or other reported values such as species
richness, basal area or stem density (Mann Whitney U test, p \textless{}
0.05). Edge influence did not extend beyond 200 m from a forest edge,
but still represented an area of \textbf{13151 ha}.

Changes in both land-use and land-cover led to concomitant changes in
AGC stocks. Recovery throughout the region was characterized for patches
of forest and plantations. Assuming high density stands, based on
previous work, this could amount to a carbon gains of 1592 Gg C across
our study area, offsetting more recent losses of approximately 1408 Gg
C. On the other hand, at the low end, if we assume a lower carbon
density of 81.8 Mg C ha\textsuperscript{-1} this would result in a total
carbon gain of 811 Gg C. Using our approach results indicate that
overall deforestation around Yangambi has resulted in a loss of
\textasciitilde{}2416 Gg C in AGC stocks.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-10}Land use land cover change statistics of forest cover around Yangambi in the central Congo Basin. The data evaluates a difference between a historical (1958) aerial photography based survey and the Hansen et al. 2013 based satellite remote sensing data. Spatial coverage statistics are provided in square kilometers (km) and hectares (ha), rounded to the nearest integer as well as Above Ground Carbon (AGC) scaled using recent survey measurements.}
\centering
\begin{tabular}[t]{lrl}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{1}{c}{AGC} \\
\cmidrule(l{3pt}r{3pt}){3-3}
  & ha & Gg C $ha\textsuperscript{-1}$\\
\midrule
Forest & 68455 & \\
-  of which edges & 13151 & \\
Regrowth > 1958 loss & 9918 & 811 - 1592\\
Loss > 2000 & 8776 & 1408\\
Loss > 1958 (permanent) & 6282 & 1008\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-11}Landscape metrics for historical and contemporary deforestation patterns. We report patch perimeter and area, their ratio and fractal dimension. Values are reported as mean $\pm$ standard deviation, across all deforestation patches.}
\centering
\begin{tabular}[t]{lllll}
\toprule
  & perimeter (m) & area (ha) & ratio ($m\textsuperscript{-1}$) & fractal dimension\\
\midrule
historical & 1451.58 $\pm$ 943.27 & 5.25 $\pm$ 5.02 & 0.03 $\pm$ 0.01 & 1.09 $\pm$ 0.04\\
contemporary & 921.74 $\pm$ 362.59 & 1.86 $\pm$ 0.75 & 0.05 $\pm$ 0.01 & 1.1 $\pm$ 0.05\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{canopy-structure-foto-texture-analysis-1}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-1}}

Visual interpretation of the scenes provide evidence that most locations
do not change dramatically with respect to canopy composition, except
for the large areas of disturbances in contemporary fallow or
young-regrowth plots. One marked difference is noted in the
mono-dominant plot 6 (Appendix Table 1). Here, the current mono-dominant
\emph{Brachystegia laurentii} is a recent development, changing the
canopy structure visibly during the last half century (\textbf{Figure
8}). The previous varied canopy structure gave way to a more dense and
uniform canopy. This is reflected in a change of the FOTO PC value from
0.19 historically to its current value of 0.54. This historical value is
similar to the mean of contemporary mono-dominant stands of
\emph{Gilbertiodendron dewevrei} with PC averaging 0.34 \(\pm\) 0.1, and
is only slightly higher than historical values for a mixed forest (0.18
\(\pm\) 0.08). The reverse pattern is seen in the contemporary PC
values. Here, the value of 0.54 exceeds those of most mono-dominant
stands (0.35 \(\pm\) 0.08), and is even further removed from the values
noted for mixed forests (0.12 \(\pm\) 0.03).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/visual_comparison_psp}

}

\caption{Visual comparison between a historical (A) and contemporary (B) permanent sampling plot. The site is currently listed as a mono-dominant Brachystegia laurenttii stand. Note the structural differences with a 'courser' canopy structure in the historical image compared to the more closed contemporary stand.}\label{fig:unnamed-chunk-12}
\end{figure}

Using only small subsets around existing permanent sampling plots we
show distinct differences between forest types, with PC values in both
historical and contemporary imagery markedly higher for the
mono-dominant forest types compared to all others (\textbf{Appendix
Figure 4}). Provided that the young-regrowth and fallow permanent
sampling plots have seen recent disturbance the Wilcoxon signed rank
test on the mixed and mono-dominant plots between the historical and
contemporary PC values did not show a significant difference (p
\textgreater{} 0.05). Similarly, no significant different using PC
values extracted from the whole scene analysis was noted (p
\textgreater{} 0.05). Any relationships between contemporary Geo-Eye
data and permanent sampling plot measurements of Above Ground Carbon,
stem density and species richness were non-significant (p \textgreater{}
0.05, Appendix \textbf{Figures 5-7}).

Furthermore, visual inspection of the scene wide analysis suggests
historical scences do not show landscape wide canopy features
(\textbf{Figure 9 A-B}), unlike the contemporary scene (\textbf{Figure 9
C-D}). Where the FOTO algorithm picks up landscape features such as
changes in texture across the contemporary Geo-Eye scene (e.g.~the river
valley as a diagonal line in \textbf{Figure 9D}), however, no
corresponding landscape patterns are found by the FOTO algorithm in the
historical orthomosaic.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/foto_maps}

}

\caption{RGB visualizations of the first three principal components of scene wide FOTO texture analysis of historical and current (Geo-Eye) imagery. Current permanent sampling plots of mono-dominant, mixed, fallow and young (edge) forest plots are marked with open triangles, open circles, open squares and crosses, respectively.}\label{fig:unnamed-chunk-13}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Finely grained spatial data sources, such as remote sensing imagery, are
rare before the satellite era (\textless{}1972). This lack of data
limits our understanding of how forest structure has varied over longer
time periods in remote areas. Long term assessment can be extended by
using large inventories of historical aerial survey data
\citep{kadmon1999, song2015, nita2018}. Despite the difficulties in
recovering this data and its limitations, such as invisible disturbances
\citep{peres2006}, remote sensing generally remains the best way to map
and quantify LULCC \citep{houghton2012}. In our study we used novel
numerical remote sensing techniques to valorize, for the first time,
historical remote sensing data in order to quantify (long term) land-use
and land-cover change and canopy structural properties in the central
Congo Basin. Despite these successes some methodological and research
considerations remain.

\hypertarget{methodological-considerations}{%
\subsection{Methodological
considerations}\label{methodological-considerations}}

\hypertarget{data-recovery-challenges}{%
\subsubsection{Data recovery
challenges}\label{data-recovery-challenges}}

In our study the archive data recovered was limited to contact prints
and therefore did not represent the true resolution of the original
negative. In addition, analogue photography clearly produces a distinct
softness compared to digital imagery (\textbf{Figure 8}). Despite
favourable nadir image acquistions \citep{verhoeven2017} image softness
combined with illumination effects between flight paths, and the
self-similar nature of vast canopy expanses
\citep{park2019, simini2010, sole1995}, limited our ability to provide
wall-to-wall coverage of the entire dataset containing 334 images. Few
man made features in the scenes also made georeferencing challenging.
Although the village of Yangambi provided a range of buildings as
(hard-edge) references, other areas within the central Congo Basin might
have fewer permanent structures and would require the use of soft-edged
landscape features (e.g.~trees, river outflows). Research has shown that
soft-edged features can help georeference scenes even when containing
few man-made features \citep{hughes2006}. Our two step georeferencing
approach resulted in a referencing accuracy of \textasciitilde{}4.7
\(\pm\) 4.3 m across reference points. However, it shoud be noted that
referencing accuracy of the final scene is less constrained toward the
edges of the scene.

\hypertarget{lulc-classification-validation}{%
\subsubsection{LULC classification \&
validation}\label{lulc-classification-validation}}

\textbf{When classifying the orthomosaic into forest and non-forest
states we favoured a deep learning supervised classification using a CNN
over manual segmentation to guarantee an ``apples-to-apples'' comparison
between the historical and the contemporary GFC forest cover map used in
our analysis. We acknowledge that both the CNN and GFC land-use and
land-cover maps use different underlying features, i.e.~spatial or
spectral data, yet attain a similarly high accuracy of up to 99\%
\citep{hansen2013}. More so, when validating our CNN classifier against
GFC data (Figure 7) for a contemporary high resolution Geo-Eye
panchromatic image we reach an accuracy of \textasciitilde{}87\%,
despite a time difference of almost 60 years. Visual inspection of the
classification data in Figure 7 suggests that the GFC map more often
than not classifies non-forest areas as forest. Actual classification
accuracy of our algorithm might therefore be higher than our reported
value.}

\hypertarget{scaling-opportunities}{%
\subsubsection{Scaling opportunities}\label{scaling-opportunities}}

\textbf{Our approach uses broadly defined homogeneous polygons to
construct a balanced dataset of synthethic landscapes. The methodology
is analoguous to the use of sparse labelling as used by
\citet{buscombe2018} and contrasts with the standard methodologies which
generally extract pixel (windows) \citep{song2015} or delineate land
cover classes \citep{nita2018} to drive a classifier or analysis. More
so, the use of heavy image augmentation during model training sidesteps
texture representation issues which affect classification of image
scenes with inconsistent illumination or sharpness \citep{hudak1998} or
ad-hoc feature engineering \citep{song2015}. The use of synthetic
landscapes allowed us to account for most, but not all, of the
variability within our orthomosaic. Our analysis has shown that despite
being trained on historical data our model could map contemporary forest
cover in remote sensing data with similar spatial and spectral
characteristics (Figure 7), suggesting that the classifier consistently
works across both space and time. We acknowledge that the use of
synthetic landscapes is limited by the available homogeneous areas to
sample from and the number of classes. Yet, the latter should not be a
constrained as previous research efforts have focussed on simple forest
loss maps \citep{hansen2013}.}

\hypertarget{research-context}{%
\subsection{Research context}\label{research-context}}

\hypertarget{long-term-changes-in-lulc-and-above-ground-carbon-1}{%
\subsubsection{Long term changes in LULC and Above Ground
Carbon}\label{long-term-changes-in-lulc-and-above-ground-carbon-1}}

Our analysis shows that the majority of deforestation around Yangambi
happened toward the late 1950's (\textbf{\textasciitilde{}16200 ha}).
Considerable regrowth has occured since the aerial survey was executed
(\textbf{\textasciitilde{}9918 ha}), and socio-economic instability
prevented further large scale forest exploitation. In particular, many
plantations have reached maturity and forest has re-established in
previously cleared or disturbed areas. The majority of this forest
recovery takes the form of isolated patches of forest but is offset by
further deforestation of previously untouched forest. Generally, the
function and structure of forests can be influenced by forest edges that
are located up to 1km away however most effects are pronounced within
the first 300 m from the edge \citep{gascon2000}. Our analysis of edge
effects on AGC has shown that the influence is negligible 200 m away
from the edge. \citet{phillips2006} have shown similar weak responses to
edge effects in the Amazon forest. Due to a lack of data on the extent
(depth) of edge effects and their influence on AGC beyond 200 m we did
not include any estimates of carbon loss or gain within these zones.
However, it must be stated that the influence of edges throughout the
landscape was not marginal as these areas would account for
\textbf{13151 ha}. Thus edges could have a substantial negative
\citep{brinck2017} or positive \citep{reinmann2017} influence on AGC.
Similarly, uncertainties in how to explicitly correct for plantations in
the landscape present a further challenge. Thus, although our estimates
are only indicative they do underscore the important influence of
landscape structure when carbon accounting. However, our findings do not
indicate that deforestation in Congo basin is declining, on the
contrary.

Over the past half century there has been a clear shift in land use in
Yangambi (Figure 5). Land use has shifted away from for example a
regular fishbone deforestation pattern that emerges when (large scale)
agricultural interests dominate the landscape \citep{arima2015}, to a
more fragmented landscape (Figure 5D). The latter former? is consistent
with historical land management at the time of the aerial survey
\citep{bustillo2018}. These regular patterns have since been reversed
because of a decrease in large scale intensive agriculture, replaced by
ad-hoc small scale subsistence farming with large perimeter to area
relationships (i.e.~ragged edges). Consequently, edge effects in the
current landscape are far more pronounced than in the historical scene.

Visual inspection of the images also suggests that reforestation within
the historically cleared areas and experimental plots is not necessarily
limited to areas far removed from more densely populated areas. For
example, large reforested areas exist close to the Congo stream and
Yangambi village itself (Fig. 5). Here, regional political components,
such as land leases and large scale ownership could have played a role
in safeguarding some of these areas for rewilding or sustainable
management \citep{arima2014, larson2011}. Despite widespread
anthropogenic influences throughout the tropics \citep{lewis2015} the
retention of these forested areas show the potential of explicit or
implicit protective policy measures (e.g.~INERA concessions,
\citet{bustillo2018}) on a multi-decadal time scale. Forest regrowth in
non-continuous areas within Yangambi could increase landscape
connectivity and help increase biodiversity \citep{vandeperre2018}.

Our analysis therefore provides an opportunity to highlight and study
those regions that have previously suffered confirmed long-term
disturbances, and those that have been restored since. Assessments of
old plantations and recovering clear-cut forests can serve as a guide to
help estimate carbon storage capacity and forest recovery rates in
managed and unmanaged conditions
\citep{gourlet-fleury2013, sader1988, achard2014} over the mid- to
long-term, in support of rewilding and general forest restoration
\citep{arima2014, larson2011, vandeperre2018}. In addition, mapping
long-term edge effects can further support research into issues such as
receding forest edges \citep{gascon2000}.

\hypertarget{canopy-structure-foto-texture-analysis-2}{%
\subsubsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-2}}

Finally, the FOTO technique used to quantify relationships between
canopy structure and forest characteristics rendered no valuable
insights of either the historical orthomosaic or recent Geo-Eye scene.
Similarly weak correlations were found previously by
\citet{solorzano2018}. In contrast, site based texture metric statistics
did show correspondence between historical and contemporary satellite
imagery. None of them were either consistent or significant. Although
visual interpretation shows distinctly different canopy structures
(Figure 5) the differences in how resolution is defined and issues
related to image quality prevented us from quantifying these further.
Unlike large scale studies by \citet{Ploton2012} we could not scale this
technique to historical data. \textbf{The successful use of our CNN
classification model on a contemporary remote sensing data does suggest
that texture can be used to consistently capture canopy properties 60
years apart. This suggests inflexibility on part of the FOTO technique
in dealing with non-standardized data.} We advise that future
valorisation efforts should preferentially work from foto negatives (if
available) to ensure optimal data quality in resolution, contrast and
sharpness.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Given the impact of tropical forest disturbances on atmospheric carbon
emissions, biodiversity and ecosystem productivity accurate long term
reporting of LULCC is an imperative. Our analysis of historical aerial
survey images (1958) of the Central Congo Basin provides a window into
the state of the forest at the start of the anthropocene. The use of a
CNN based LULC classifier, using synthetic landscapes based image
augmentation, provides a robust semi-supervised solution which scales
across space and time, even for image scenes with inconsistent
illumination or sharpness. Combined with contemporary remote sensing
data we have shown that historical aerial survey data can be used to
quantify long-term changes in LULC and AGC. We showed a shift from
previously highly structured industrial deforestation of large areas
\textbf{for plantation purposes}, to discrete smallholder clearing for
farming, increasing landscape fragmentation but also opportunties for
substantial regrowth. Efforts to quantify canopy texture features and
their link to AGC had limited to no success. Our analysis provides
insights into the rate at which deforestation and reforestation has
taken place over a multi-decadal scale in the central Congo basin. As
such, it provides a useful historical context while interpreting past
and ongoing forest research in the area.

\hypertarget{additional-information-and-declarations}{%
\section{Additional Information and
Declarations}\label{additional-information-and-declarations}}

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

Hufkens et al. (2019). A curated dataset of aerial survey images over
the central Congo Basin, 1958. Zenodo:
\url{https://doi.org/10.5281/zenodo.3547767}. All data not included in
the latter repository can be found bundled with the analysis code as
listed below. Proprietary datasets (i.e.~Geo-Eye data) are not shared,
but purchase order numbers allow for acquisition of these datasets to
ensure reproducibility.

\hypertarget{code-availability}{%
\subsection{Code availability}\label{code-availability}}

All analysis code is available as an R / python \citep{rcoreteam2019}
projects (\url{https://khufkens.github.io/orthodrc} \&
\url{https://khufkens.github.io/orthodrc_cnn/}). The analysis relied
heavily on the `raster' \citep{hijmans2019}, `RStoolbox'
\citep{leutner2019}, and `landscapemetrics' \citep{hesselbarth2019}
packages, while post-processing and plotting was facilitated by the
`tidyverse' ecosystem \citep{wickham2017}, `ggthemes'
\citep{arnold2019}, `scales' \citep{wickham2018} and `cowplot'
\citep{wilke2019}. Additional plotting elements were formatted or
provided by `sf' \citep{pebesma2018} and `rnaturalearth'
\citep{south2017} packages, respectively. I'm grateful for the
contributions to the scientific community by the developers of these
packages.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
%
% % Only for the journal Methods and Protocols:
% % If you wish to submit a video article, please do so with any other supplementary material.
% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This research was supported through the Belgian Science Policy office
COBECORE project (BELSPO; grant BR/175/A3/COBECORE). KH acknowledges
funding from the European Union Marie Skłodowska-Curie Action (project
number 797668). I'm grateful for the support of Thales de Haulleville
volunteering his time in scanning the images.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{K.H. conceived and designed the study, analyzed the data, prepared
figures, tables and authored the final draft of the manuscript. T.d.H.
scanned all image data. E.K. and T.d.H. provided plot based AGC
estimates. T.D., K.J., E.K, H.B., P.S., F.V.,S.M., M.A., J.V.D.B., H.V
and L.W. reviewed the final manuscript.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest. The founding sponsors had
no role in the design of the study; in the collection, analyses, or
interpretation of data; in the writing of the manuscript, an in the
decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{bibliography.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}