### 2020-02-06 manuscript/main.tex

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[remote sensing,article,submit,moreauthors,pdftex]{mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

\usepackage{longtable}

% see https://stackoverflow.com/a/47122900

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Historical aerial surveys map long-term changes of forest cover and
structure in the central Congo Basin}

% Authors, for the paper (add full first names)
\Author{Koen Hufkens$^{1,2,*}$\href{https://orcid.org/0000-0002-5070-8109}{\orcidicon}, Thalès de Haulleville$^{3}$, Elizabeth Kearsley$^{1}$, Kim Jacobsen$^{1,3}$, Hans Beeckman$^{3}$, Piet Stoffelen$^{4}$, Filip Vandelook$^{4}$, Sofie Meeus$^{4}$, Michael Amara$^{5}$, Leen Van Hirtum$^{5}$, Jan Van den Bulcke$^{1}$, Hans Verbeeck$^{1}$, Lisa Wingate$^{2}$}

% Authors, for metadata in PDF
\AuthorNames{Koen Hufkens, Thalès de Haulleville, Elizabeth Kearsley, Kim Jacobsen, Hans Beeckman, Piet Stoffelen, Filip Vandelook, Sofie Meeus, Michael Amara, Leen Van Hirtum, Jan Van den Bulcke, Hans Verbeeck, Lisa Wingate}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Ghent University, Ghent, Belgium; \\
$^{2}$ \quad INRA, UMR ISPA, Villenave d'Ornon, France; \\
$^{3}$ \quad Royal Museum for Central Africa, Tervuren, Belgium; \\
$^{4}$ \quad Botanic Garden Meise, Meise, Belgium; \\
$^{5}$ \quad National Archives of Belgium, Brussels, Belgium; \\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:koen.hufkens@gmail.com}{\nolinkurl{koen.hufkens@gmail.com}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Given the impact of tropical forest disturbances on atmospheric carbon
emissions, biodiversity and ecosystem productivity, accurate long term
reporting of Land-Use and Land-Cover (LULC) change in the pre-satellite
era (\textless{} 1972) is an imperative. Here, we used a combination of
historical (1958) aerial photography and contemporary remote sensing
data to map long-term changes in the extent and structure of the
tropical forest surrounding Yangambi (DR Congo), in the central Congo
Basin. Our study leveraged Structure-from-Motion and a Convolutional
Neural Network based LULC classifier, using synthetic landscapes based
image augmentation, to map historical forest cover across a large
orthomosaic (\textasciitilde{}82 800 ha) geo-referenced to
\textasciitilde{}4.7 \(\pm\) 4.3 m at sub meter resolution. A comparison
with contemporary LULC data showed a shift from previously highly
regular industrial deforestation of large areas, to discrete smallholder
farming clearing, increasing landscape fragmentation but also providing
opportunties for substantial forest regrowth. We estimated above ground
carbon gains through reforestation to range from 811 to 1592 Gg C,
partially offseting historical deforestation (2416 Gg C), in our study
area. Efforts to quantify long term canopy texture changes and their
link to above ground carbon had limited to no success. Our analysis
provides methods and insights into key spatial and temporal patterns of
deforestation and reforestation at a multi-decadal scale, providing a
historical context for past and ongoing forest research in the area.}

% Keywords
\keyword{Aerial survey, data recovery, CNN, Deep Learning, SfM, Congo Basin}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{wrapfig} \usepackage{graphics} \usepackage{setspace} \doublespacing \usepackage{gensymb} \usepackage{booktabs}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Tropical ecosystem services are severely impacted by deforestation and
forest degradation \citep{hansen2013, houghton2012, tyukavina2015}. Not
only does tropical forest Land-Use and Land-Cover Change (LULCC)
constitute 10 to 15\% of the total global carbon emissions
\citep{vanderwerf2009}, changes in forest fragmentation affect the
forest structure and function
\citep{fauset2017, brinck2017, didham1999}. Strong fragmentation effects
decrease the number of large trees along forest edges
\citep{laurance2000, magnago2015}, while species composition and
biodiversity are equally negatively affected
\citep{poorter2006, barlow2016, vandeperre2018}. Estimates show that
31\% of carbon emissions are caused by edge effects alone
\citep{brinck2017}.

Accurate estimates of LULCC and forest canopy structure are therefore
imperative to estimate carbon emissions and other ecosystem services
\citep{houghton2012, hansen2013}. Remote sensing products have been key
inputs in LULCC assessments as they provide accurate spatial information
to help estimate carbon emissions \citep{mitchard2018, hansen2013}. More
so, high resolution aerial images provide scientists tools to monitor
forest extent, structure and carbon emissions as canopy texture is
linked to above ground biomass
\citep{Ploton2012, Couteron2005, Barbier2010}. Yet, most of these
estimates are limited in time to recent decades
\citep{defries2002, houghton2012, hansen2013, achard2014}.

Historical estimates of Land-Use and Land-Cover (LULC), in the
pre-satellite era (\textless{} 1972), exist but generally rely on
non-spatially explicit data (i.e socio-economic data)
\citep{houghton2012, ramankutty1999, defries2002, sader1988}. Efforts
have been made to use other geo-spatial data sources such as historical
maps \citep{willcock2016}, declassified CORONA satellite surveillance
data across the US and central Brazil \citep{song2015} as well as aerial
surveys in post World War II Germany \citep{nita2018}. Survey data
across the African continent is less common, inaccessible or both. Some
studies do exist, as \citet{buitenwerf2012} and \citet{hudak1998} used
aerial survey images to map vegetation changes in South African
savannas, whilst \citet{frankl2015} and \citet{nyssen2016} mapped the
Ethiopian highlands of 1930s.

Across the central Congo Basin most of these historical images were
collected within the context of national cartographic efforts by the
``Institut Géographique du Congo Belge'' in Kinshasa (then
Léopoldville), DR Congo. Despite the existence of large archives of
aerial survey imagery of African rainforest (Figure 1, Appendix Figure
3), as of yet, no studies have valorized these data. The lack of a
consistent valorization effort is unfortunate as the African rainforest
is the second largest on Earth and covers \textasciitilde{}630 million
ha, representing up to 66 Pg of carbon storage \citep{Lewis2009b}, and
currently loses forest at an increasing pace \citep{butsic2015}. Given
the impact of LULCC on the structure and functioning of central African
tropical forests, and their influence on both carbon dynamics
\citep{Kearsley2013} and biodiversity \citep{vandeperre2018}, accurate
long term reporting of historical forest cover warrants more attention
\citep{willcock2016}.

Here, we use a combination of historical aerial photography (1958) and
contemporary remote sensing data (2000-2018) to map long-term changes in
the extent and structure of the tropical forest surrounding Yangambi (DR
Congo), in the central Congo Basin, effectively linking the start of the
anthrophocene \citep{lewis2015} with current assessments. Yangambi was,
and remains, a focal center of forest and agricultural research and
development in the central Congo Basin. Past research in the region
allows for thorough assessment of LULCC from a multi-disciplinary point
of view, confronting us with complex deforestation and land-use
patterns.

We leverage Structure-from-Motion to generate a large orthomosaic of
historical imagery and develop a Convolutional Neural Network based
forest cover mapping approach, based upon a semi-supervised generated
dataset extensively leveraging data augmentation. Our methodology aims
to provide a historical insight into important LULCC spatial patterns in
Yangambi, such as fragmentation and edge complexity. We further
contextualize the influence of changes in the forest's life history on
past and current research into Above Ground Carbon (AGC) storage
\citep{Kearsley2013} and biodiversity \citep{vandeperre2018} in the
central Congo Basin. Our fast scalable mapping approach for historical
aerial survey data, using limited supervised input, would further
support long-term land-use and land-cover change analysis across the
central Congo Basin.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{historical-data-acquisition}{%
\subsection{Historical data
acquisition}\label{historical-data-acquisition}}

Data for the central Congo Basin region, surrounding Kisangani, were
collected in several flights during the dry season of 1958 and 1959
(from 8/01/1958 to 20/02/1958 and from 28/12/1958 to 9/01/1959
respectively, see Appendix Figure 1 - 2 and Appendix Table 1) to
generate topographic maps of the area, supervised by the ``Institut
Géographique du Congo Belge'' in Kinshasa, DR Congo (then Léopoldville).
Black-and-White panchromatic images (0.4 - 0.9 \(\mu\)m) were gathered
along flight paths running mostly from west to east, between 09:00 and
11:00 local time. Along a flight path continuous images were taken using
a Wild Heerburgg RC5a (currently Leica Geosystems) with an Aviogon lens
assembly (114.83mm / f 5.6, with a 90\degree view angle) resulting in
square photo negatives of 180 by 180 mm. Flights were flown at an
average absolute altitude of \textasciitilde{}5200 m above sea level,
covering roughly 18 530 km\textsuperscript{2} at an approximate scale of
1/40 000. The use of the integrated autograph system ensured timely
acquisition of pictures with a precise overlap (\textasciitilde{}1/3)
between images. This large overlap between images together with flight
parameters would allow post-processing, using stereographs, to create
accurate topographic maps. Original data from this campaign are stored
in the \href{https://www.africamuseum.be/en}{Royal Museum for Central
Africa} in Tervuren, Belgium (Figure 1, Appendix Figure 3).

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{./figures/print_examples}

}

\caption{A box of historical aerial photographs (left) and a single aerial photograph (right) showing part of the Congo river. Note the meta-data provided in the right side margin of the image, such as acquisition time and flight height.}\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{site-selection}{%
\subsection{Site selection}\label{site-selection}}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{./figures/flight_paths}

}

\caption{Overview of the historical flight paths during aerial photo acquisition and ancillary data used in this study. The bounding box of the orthomosaic data presented in this study is shown as a rectangle (23x36 km). The outline of a recent high-resolution Geo-eye panchromatic image is shown as a dashed dark grey rectangle (10x10 km). The location of various permanent sampling plots are shown as x, +, and open squares and triangles for the mixed, mono-dominant and edge plots respectively. The grey polygon delineates the current day Yangambi Man and Biosphere reserve. The inset, top right, situates the greater Yangabmi region (white rectangle) with the DR Congo. The full flight plan and details are shown in Appendix Figure 1 and 2}\label{fig:unnamed-chunk-2}
\end{figure}

We prioritised flight paths and images that contained current day
permanent sampling plots, larger protected areas, and past agricultural
and forest research facilities (Figure 2). This selection provides a
comprehensive mapping of the Yangambi area and the life history of the
forest surrounding it. Thereafter, we selected flight paths 1 through 11
for digitization. From this larger dataset of 334 images we selected 74
survey images for orthomosaic compositing and further analysis. All the
selected images stem from the flight campaign made during January and
February of 1958. The area includes the Yangambi village, 20
contemporary permanent sampling plots \citep{Kearsley2013}, past and
present agricultural experimental plots \citep{bauters2015} and large
sections of the Yangambi
\href{http://www.unesco.org/new/en/natural-sciences/environment/ecological-sciences/biosphere-reserves/africa/democratic-republic-of-the-congo/yangambi/}{UNESCO
Man and Biosphere reserve} surrounding to the west and east of the
village. Although not formally mosaicked we provided a full dataset of
pre-processed images using the cropping and normalization routines
described below. The latter data was not used in subsequent LULCC
analysis, but has been archived and made available to the public
separately (see code \& data availability statement below).

\hypertarget{digitization-and-data-processing}{%
\subsection{Digitization and data
processing}\label{digitization-and-data-processing}}

All selected images, covering the Yangambi area, were contact prints as
original negatives of the prints were not available. Images were scanned
at a resolution exceeding their original resolution (or grain) at the
maximal physical resolution of an Epson A3 flatbed scanner (i.e.~2400
dpi or 160 MP per image) and saved as lossless tiff images. Data were
normalized using contrast limited histogram equalization
\citep{zuiderveld1994} with a window size of 32 and a clip limit of 1.5.
Fiduciary marks were used to rectify and downsample the images into
square 7700x7700 pixel images (\textasciitilde{}1200 dpi, 81 MP). This
resulted in a dataset with digital images at a resolution that remained
above the visible grain of the photographs, whilst the reduced image
size facilitated easier file handling and processing speed.

Data was processed into a georeferenced orthomosaic using a
Structure-from-motion (SfM, \citet{ullman1979}) approach implemented in
\href{http://www.agisoft.com}{Agisoft Metashape} version 1.5.2 (Agisoft
LLC, St.~Petersburg, Russia). An orthomosaic corrects remote sensing
data to represent a perfectly downward looking image, free from
perspective distortions due to topography and camera tilt. Using the SfM
technique features, areas in images with a large degree of similarity,
are matched across various images to reconstruct a three dimensional
scene (topography) from two-dimensional image sequences. During the SfM
analysis we masked most clouds, glare or large water bodies such as the
Congo river.

We calculated the orthomosaic using a low resolution point cloud and
digital elevation map (DEM). Additional ground control points were
provided to assist in the referencing of image and constrain the
optimization routine used in the SfM algorithm. Ground control points
consisted of permanent structures which could be verified in both old
and new aerial imagery (i.e.~ESRI World Imagery) and consisted of corner
points of build structures (e.g.~a building, bridge or swimming pool
etc.). Although most clouds were removed during the SfM routine some
were retained to provide sufficient SfM tie points to maximize
continuous forest coverage in the final orthomosaic. The final scene was
cropped, to provide consistent wall-to-wall coverage of the
reconstructed scene. The orthomosaic was exported as a geotiff for
further georeferencing in QGIS \citep{qgisdevelopmentteam2018} using the
georeferencer plugin (version 3.1.9) and additional ESRI World Imagery
high resolution reference data. We used 3\textsuperscript{rd} degree
polynomial and 16 ground control points to correct the final image.
Ground control points, raw image data and final processed image are
provided in addition to measures of uncertainty such as mean and median
error across all ground control points. All subsequent analysis are
executed on the final geo-referenced orthomosaic or subsets of it.

\hypertarget{land-use-and-land-cover-change}{%
\subsection{Land-Use and Land-Cover
Change}\label{land-use-and-land-cover-change}}

\hypertarget{classifying-land-use-and-land-cover}{%
\subsubsection{Classifying Land-Use and
Land-Cover}\label{classifying-land-use-and-land-cover}}

\hypertarget{model-training}{%
\paragraph{\texorpdfstring{\emph{Model
training}}{Model training}}\label{model-training}}

We automatically delineated all natural forest in the historical data,
thus excluding tree plantations, thinned or deteriorated forest stands
which showed visible canopy cover loss, fields and buildings. We used
the Unet Convolutional Neural Net (CNN, \citet{ronneberger2015})
architecture implemented in Keras \citep{chollet2015} with an
efficientnetb3 pre-processing backbone \citep{yakubovskiy2019} running
on TensorFlow \citep{martinabadi2015} to train a binary classifier
(i.e.~forest or non-forested). Training data were collected from the
orthomosaic by randomly selecting 513 pixel square tiles from locations
within homogeneous forested or non-forested polygons in the historical
orthomosaic (Figure 5). Separate polygons were selected for training,
testing and validation purposes. Validation polygons were sampled 300
times, while both testing and validation polygons were sampled at 100
random locations. Tiles extracted from locations close to the polygon
border at times contained mixed cover types. Tiles with mixed cover
types were removed from the list of source tiles (Table 1). Homogeneous
source tiles were combined in synthetic landscapes using a random
gaussian field based binary mask (Figure 3). We generated 5000 synthetic
landscapes (balancing forest and non-forest classes) for training, while
500 landscapes were generated for both the validation and the testing
datasets for a total of 6000 synthetic landscapes. In order to limit
stitch line misclassifications, along the seams of mosaicked images, we
created synthetic landscapes with different forest tiles to mimick
forest texture transitions. We applied this technique to 10\% of the
generated synthetic landscapes (across training, validation and testing
data).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Number of source tiles used for the generation of synthetic landscapes.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  & training & testing & validation\\
\midrule
forest & 300 & 100 & 100\\
non-forest & 294 & 92 & 84\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/synthetic_landscape}

}

\caption{An example synthetic landscape, combining homogeneous forest and non-forest images into a patchy landscape using random gaussian field based masks. The left panel shows a combined synthetic landscape, while the right panel shows the corresponding forest (black) and non-forest (white) labels.}\label{fig:unnamed-chunk-4}
\end{figure}

The CNN model was trained for 100 epochs with a batch size of 30 using
Adam optimization \citep{kingma2017}, maximizing the
Intersect-over-Union (IoU) using Sørensen--Dice \citep{dice1945} and
categorical cross-entropy loss functions. Data augmentation included
random cropping to 320 pixel squares, random orientation, scaling,
perspective, contrast and brightness shifts and image blurring. The
optimized model was used to classify the complete orthomosaic using a
moving window approach with a step size of 110 pixels and a majority
vote (\textgreater{} 50\% agreement) across overlapping areas to limit
segmentation edge effects. In addition, we provide raw pixel level
classification agreement data for quality control purposes (see Data
availability below). We refer to Figure 4 for a synoptic overview of the
full deep learning workflow.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/cnn_diagram}

}

\caption{A diagram of the deep learning workflow followed in training a binary forest / non-forest cover convolutional neural net U-Net model to generate our forest cover map.}\label{fig:unnamed-chunk-5}
\end{figure}

\hypertarget{model-validation}{%
\paragraph{\texorpdfstring{\emph{Model
validation}}{Model validation}}\label{model-validation}}

We report the CNN accuracy based upon the IoU of our out-of-sample
validation dataset of synthetic landscapes. In addition, we report
confusion matrices for all pixels across the homogeneous validation
polygons, as well as the training and testing polygons (see Figure 5).
Furthermore, we used the first acquisition of a recent pan-chromatic
Geo-Eye 1 stereo pair (Geo-Eye, Thornton, Colorado, U.S.A., order
737537, 2011-11-11 08:55 GMT or 09:55 local time) to classify and assess
the robustness of the CNN algorithm on contemporary remote sensing
imagery with similar spectral and spatial characteristics. We used the
Global Forest Change version 1.6 (GFC, tile 10N-020E) \citep{hansen2013}
map data, capturing deforestation up to 2011, to quantify accuracy on
downsampled CNN Geo-Eye classification results. Once more, we report the
confusion matrix between the GFC and CNN derived forest cover maps,
masking clouds and cloud shadows. To summarize confusion matrices we
report accuracy as:

\(Accuracy = \frac{(TP + TN)}{(TP + TN + FP + FN)}\)

in which TP, TN, FP, FN are True Positive, True Negative, False Positive
and False Negative, respectivelly.

\hypertarget{characterizing-long-term-change}{%
\subsubsection{Characterizing long term
change}\label{characterizing-long-term-change}}

To map long term LULCC in the Yangambi region we used the contemporary
Global Forest Change version 1.6 (GFC, tile 10N-020E) (lossyear) map
data \citep{hansen2013}. Using the GFC data we calculated the latest
state of the forest with respect to the conditions at the start of 1958,
60 years earlier. In our analysis we only included GFC pixels which
recorded no forest loss throughout the whole 2000 - 2018 period. Forest
loss in the context of GFC is defined as ``a stand-replacement
disturbance, or a change from a forest to non-forest state''. As such,
locations which would see reforestation or deforestation between 2000
and 2018 would be marked as non-forest (i.e.~disturbed). As the
resolution of the historical forest classification exceeds that of the
GFC map we downsampled our historical forest cover data to 30 m GFC
resolution using a nearest neighbour approach. We masked out all water
bodies using the Global Forest Change survey data mask layer, and
limited the analysis to the right bank of the Congo river. We provide
summary statistics of historical and contemporary deforestation and
reforestation. We map permanent deforestation after 1958, reforestation
after 1958, recent deforestation and long term (stable) forest cover.
All references to changes over time in the context of our analysis
explicitly compare the historical and contemporary periods from hereon
forth.

\hypertarget{landscape-fragmentation-above-ground-carbon-estimates}{%
\subsubsection{Landscape fragmentation \& Above Ground Carbon
estimates}\label{landscape-fragmentation-above-ground-carbon-estimates}}

To quantify changes in the structure of forest cover and its
disturbances we used spatial landscape pattern analysis
(i.e.~fragmentation) metrics \citep{dale1999}. Landscape metrics provide
a mathematical framework for the analysis of discrete land‐cover classes
and allows us to capture their composition and configuration. These
metrics are therefore commonly used to compare how landscapes change
over time \citep{hesselbarth2019}. In particular, fractals provide a way
to quantify complex natural landscapes, including their self-similarity
across scales \citep{Li2000, Mandelbrot1977}. We report the ratio of
edge to area and the fractal dimension to quantify landscape complexity
of forest disturbances. A fractal dimension closer to 2 suggest a more
complex (fragmented) landscape.

Statistics were calculated for all forest disturbance patches larger
than 1 ha and smaller than the 95\textsuperscript{th} percentile of the
patch size distribution using the R package ``landscapemetrics''
\citep{hesselbarth2019}. We provide mean and standard deviation on edge,
area, their ratio and fractal dimension for both the historical and
contemporary \citet{hansen2013} forest cover maps.

We estimated above ground carbon (AGC) losses and gains, due to
deforestation and reforestation, using plot based averages of recent
inventory data at permanent sampling plots in the area (Figure 2). We
refer to \citet{Kearsley2013} for the survey method and allometric
relations used to scale the survey data. Unlike standard square 1 ha
plots, edge plots (163.03 \(\pm\) 19.39 Mg C ha\textsuperscript{-1}, N =
5) were set back 200 m from forest edges and were 50x200 m, with the 50
m side of the plot along the forest edge and continuing 200 m into the
forest (Appendix Table 2). We further confirmed that forest edge plots,
as compared to mixed forest plots (160.48 \(\pm\) 23.84 Mg C
ha\textsuperscript{-1}, N = 8, see Appendix Table 3 for all forest
types), did not show a significantly different AGC (Mann Whitney U test,
p \textless{} 0.05). Thus it was not necessary to explicitly quantify
changes in AGC caused by edge effects. Moreover, we used the mean value,
and its uncertainty (i.e.~standard deviation), of the mixed forest as
representative for potential AGC losses. Despite the challenges inherent
in quantifying AGC for forest edges we mapped the total extent of the
edges in the contemporary landscape. To align our landscape analysis
with exploratory analysis of the survey data we used a buffer of 200 m
to estimate the extent of forest edges and patches, up to the location
of forest edge plots.

Surveys of old plantations show a large variation in AGC, depending on
age and the crop type. For example, the AGC values varied from 86.55 to
168.67 Mg C ha\textsuperscript{-1}, for \emph{Elaeis guineensis} (oil
palm) and \emph{Hevea brasiliensis} (rubber tree) plots respectively
(\citet{bustillo2018}, personal communications). These higher values are
in line with the mixed plot AGC estimates (160.48 \(\pm\) 23.84 Mg C
ha\textsuperscript{-1}, N = 8) in the area, while the palm plantations
resemble old-regrowth values (81.87 Mg C ha\textsuperscript{-1}, N = 1).
To quantify AGC in reforested areas we therefore use both AGC estimates
of old-regrowth and mixed forest, as lower and upper bounds. We did not
have sufficient data to account for individual changes in AGC across
plantations.

\hypertarget{canopy-structure-foto-texture-analysis}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis}}

We compared the structure of the canopy both visually and using Fourier
Transform Textural Ordination (FOTO, \citet{couteron2002}). FOTO uses a
principal component analysis (PCA) on radially averaged 2D Fourier
spectra to characterize canopy (image) texture. The FOTO technique was
first described by \citet{couteron2002} to quantify canopy stucture in
relation to biomass and biodiversity, and can be used across multiple
scenes using normalization \citep{Barbier2010}.

We used the standard FOTO methodology with fixed zones, instead of the
moving window approach. The window size was set to the same size (187
pixels or \textasciitilde{}150 m) as used in the moving window analysis
above. To account for illumination differences between the two scenes we
applied histogram matching. No global normalization was applied, as the
scene was processed as a whole. PC values from this analysis for all
permanent sampling plots in both image scences were extracted using a
buffer with a radius of 50 m.

For both site based and scene analysis we correlated PC values with
permanent sample plot inventory data such as stem density, above ground
biomass and tree species richness (Appendix Table 2-3). Additional
comparisons are made between contemporary Geo-Eye data and the
historical orthomosaic derived PC values. Due to the few available
permanent sampling plots in both scenes we used a non-parametric paired
signed rank (Wilcoxon) test \citep{wilcoxon1945} to determine
differences between the PC values of the Geo-Eye and historical
orthomosaic image scenes across mono-dominant and mixed forest types. In
all analysis, mono-dominant site 4 was removed from the analysis due to
cloud contamination.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{orthomosaic-construction}{%
\subsection{Orthomosaic construction}\label{orthomosaic-construction}}

Our analysis provides a first spatially explicity historical composite
of aerial survey images in support of mapping land-use and land-cover
within the Congo Basin. The use of high resolution historical images
combined with SfM image processing techniques allowed us to mosaic old
imagery across a large extent. The final orthomosaic composition of the
Yangambi region provided an image scene covering approximately 733
million pixels across \textasciitilde{}82 800 ha with a resolution of
0.88 m / pixel (or \textasciitilde{}23x36 km, Figure 2). The overall
spatial accuracy of the SfM orthomosaic composition using the sparse
cloud DEM (with a resolution of 45.8 m/pixel) was limited to
approximately 23 m. Further georeferencing outside the SfM workflow
reduced the mean error at the ground control points to 5.3 \(\pm\) 4.9
px (\textasciitilde{}4.7 \(\pm\) 4.3 m), with a median error of 2.9 px
(2.6 m). The orthomosaic served as input for all subsequent LULCC
analysis with all derived maps provided with the manuscript repository
(see data \& code availability statements below).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/orthomosaic_maps}

}

\caption{Overview of the final orthomosaic of the greater Yangambi region (A) as well as the forest cover classification uncertainty (B) used to generate the final Land Cover map (see Figure 6). Homogeneous polygons used in the generation of the synthetic landscape for Convolutional Neural Network training, testing and validation are marked as dashed and full lines for forest and non-forest regions, respectively. Training, testing and validation regions are denoted in panel B in green, red and orange, respectivelly. Black polygon outlines denote cloud and image stitch line regions which were manually excluded from analysis, but retained in calculation of validation statistics (see Table 2).}\label{fig:unnamed-chunk-6}
\end{figure}

\hypertarget{land-use-and-land-cover-classification}{%
\subsection{Land-use and Land-Cover
Classification}\label{land-use-and-land-cover-classification}}

\hypertarget{cnn-model-validation}{%
\subsubsection{CNN model validation}\label{cnn-model-validation}}

The CNN deep learning classifier reached an Intersection-over-Union
accuracy of 97\% on the detection of disturbed forest in the
out-of-sample (validation) synthetic landscape data. Using all pixels
within the validation polygons (Figure 5) showed a similar accuracy
value of \textasciitilde{}98\%. Using all polygons across the scene,
including those used in the generation of testing and training synthetic
landscapes, increased the accuracy to \textasciitilde{}99\% (Table 2). A
comparison with recent pan-chromatic Geo-Eye data shows good agreement,
with an accuracy of \textasciitilde{}87\% across all pixels, between the
landsat based GFC data and downscaled CNN results (Table 2 and Figure
7).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-7}Confusion matrix showing \% agreement between forest / non-forest classes using a Convolutional Neural Network (CNN) across previously selected homogenous areas. In addition, overall accuracy is reported for each confusion matrix.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
\multicolumn{1}{c}{CNN} & \multicolumn{2}{c}{Geo-Eye (full scene)} & \multicolumn{2}{c}{Aerial Survey (all polygons)} & \multicolumn{2}{c}{Aerial Survey (validation polygons)} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
  & non-forest & forest & non-forest & forest & non-forest & forest\\
\midrule
non-forest & 10.17 & 8.55 & 97.4 & 0.19 & 98.59 & 0.0\\
forest & 3.75 & 77.52 & 2.6 & 99.81 & 1.41 & 100.0\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Accuracy}}\\
\hspace{1em} &  & 87.70 &  & 98.61 &  & 99.3\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/forest_cover_map}

}

\caption{Overview of the final Land Use Land Cover Change map, a detailed inset of both the underlying orthomosaic (B) and the derived land use land cover change map displayed as the difference between the Convolutional Neural Network based classification orthomosaic and the recent Landsat based forest cover map by Hansen et al. 2013 (C).}\label{fig:unnamed-chunk-8}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/visual_comparison_classifiers}

}

\caption{Convolutional Neural Network (CNN) based forest cover classification results (B) as run on a recent (2011) Geo-Eye panchromatic image (A). We show the difference between the Convolutional Neural Network based classification and a recent Landsat based forest cover map by Hansen et al. 2013. Full white outlines denote cloud contamination, the dashed rectangle shows a location where the CNN outperforms the Landsat based forest classification.}\label{fig:unnamed-chunk-9}
\end{figure}

\hypertarget{long-term-changes-in-lulc-and-above-ground-carbon}{%
\subsection{Long term changes in LULC and Above Ground
Carbon}\label{long-term-changes-in-lulc-and-above-ground-carbon}}

Scaling our classifier to the whole historical orthomosaic we detected
\textasciitilde{}16 200 ha (or \textasciitilde{}20\% of the scene) of
disturbed forests. A large fraction of the disturbed area was restored
in the period between the two acquisition periods. In total, 9918 ha, or
little over half of the affected forest was restored (Figure 5C-D, dark
blue). Recent deforested areas, as registered through satellite remote
sensing (\textgreater{} 2000), approximate 8776 ha (Table 3, Figure 6 -
light green).

Recent deforestation follows a distinctly different pattern compared to
historical patterns. Historical deforestation showed a classical
fishbone pattern for forest clearing with very sharp edges, while
current patterns are patchy and ad-hoc (figure 6C, dark blue and green
colours respectively). These differences are reflected in the analysis
of landscape metrics of deforestation. Between the historical and
contemporary LULCC maps we see an increase in small disturbances, as
indicated by the decreasing area of the mean patch size, down to
\textasciitilde{}1.86 \(\pm\) 0.75 ha from \textasciitilde{}5.25 \(\pm\)
5.02 ha historically. Perimeter lengths were longer historically, at
1451 \(\pm\) 943 m, compared to contemporary landscapes
\textasciitilde{}921 \(\pm\) 362 m (Table 4). This shift in perimeter
area ratio led to a similar change in the fractal index, slightly
increasing in value to 1.1 \(\pm\) 0.05 from 1.09 \(\pm\) 0.04 over
time. Values closer to a fractal index of 2 suggest a more complex
(fragmented) landscape.

A comparison of forest edge plots with mixed forest plots showed no
signficant difference in AGC, or other reported values such as species
richness, basal area or stem density (Mann Whitney U test, p \textless{}
0.05). Edge influence did not extend beyond 200 m from a forest edge,
but still represented an area of 13 151 ha (Table 3).

Changes in both land-use and land-cover led to concomitant changes in
AGC stocks. Recovery throughout the region was characterized for patches
of forest and plantations. Assuming high density stands, based on
previous work, this could amount to a carbon gains of up to 1592 Gg C
across our study area, offsetting more recent losses of approximately
1408 \(\pm\) 209 Gg C. On the other hand, at the low end, if we assume a
lower carbon density of 81.8 Mg C ha\textsuperscript{-1} this would
result in a total carbon gain of 811 Gg C. Using our approach results
indicate that overall deforestation around Yangambi has resulted in a
loss of \textasciitilde{}2416 \(\pm\) 358 Gg C in AGC stocks.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-10}Land use land cover change statistics of forest cover around Yangambi in the central Congo Basin. The data evaluates a difference between a historical (1958) aerial photography based survey and the Hansen et al. 2013 based satellite remote sensing data. Spatial coverage statistics are provided hectares (ha), rounded to the nearest integer as well as Above Ground Carbon (AGC) scaled using recent survey measurements.}
\centering
\begin{tabular}[t]{lrl}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{1}{c}{AGC} \\
\cmidrule(l{3pt}r{3pt}){3-3}
  & ha & Gg C\\
\midrule
Non-forest cover (1958) & 16200 & \\
Forest cover & 68455 & \\
-  of which forest edges & 13151 & \\
reforestation >1958 & 9918 & 811 - 1592\\
deforestation >2000 & 8776 & 1408 $\pm$ 209\\
\addlinespace
deforestation >1958 (permanent) & 6282 & 1008 $\pm$ 150\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-11}Landscape metrics for historical and contemporary deforestation patterns. We report patch perimeter and area, their ratio and fractal dimension. Values are reported as mean $\pm$ standard deviation, across all deforestation patches.}
\centering
\begin{tabular}[t]{lllll}
\toprule
  & perimeter (m) & area (ha) & ratio ($m\textsuperscript{-1}$) & fractal dimension\\
\midrule
historical & 1451.58 $\pm$ 943.27 & 5.25 $\pm$ 5.02 & 0.03 $\pm$ 0.01 & 1.09 $\pm$ 0.04\\
contemporary & 921.74 $\pm$ 362.59 & 1.86 $\pm$ 0.75 & 0.05 $\pm$ 0.01 & 1.1 $\pm$ 0.05\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{canopy-structure-foto-texture-analysis-1}{%
\subsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-1}}

Visual interpretation of the scenes provide evidence that most locations
do not change dramatically with respect to canopy composition, except
for the large areas of disturbances in contemporary fallow or
young-regrowth plots. One marked difference is noted in the
mono-dominant plot 6 (Appendix Table 2). Here, the current mono-dominant
\emph{Brachystegia laurentii} is a recent development, changing the
canopy structure visibly during the last half century (Figure 8). The
previous varied canopy structure gave way to a more dense and uniform
canopy. This is reflected in a change of the FOTO PC value from 0.19
historically to its current value of 0.54 (Figure 9). This historical
value is similar to the mean of contemporary mono-dominant stands of
\emph{Gilbertiodendron dewevrei} with PC averaging 0.34 \(\pm\) 0.1, and
is only slightly higher than historical values for a mixed forest (0.18
\(\pm\) 0.08, Figure 9). The reverse pattern is seen in the contemporary
PC values. Here, the value of 0.54 exceeds those of most mono-dominant
stands (0.35 \(\pm\) 0.08), and is even further removed from the values
noted for mixed forests (0.12 \(\pm\) 0.03, Figure 9).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/visual_comparison_psp}

}

\caption{Visual comparison between a historical (A) and contemporary (B) permanent sampling plot. The site is currently listed as a mono-dominant Brachystegia laurenttii stand. Note the structural differences with a 'courser' canopy structure in the historical image compared to the more closed contemporary stand.}\label{fig:unnamed-chunk-12}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{./figures/foto_bplot_psp}

}

\caption{Boxplots comparing the first principal component (PC1) of a site based FOTO analysis across different forest types for both contemporary (Geo-Eye) and historical orthomosaic data.}\label{fig:unnamed-chunk-13}
\end{figure}

Using only small subsets around existing permanent sampling plots we
show distinct differences between forest types, with PC values in both
historical and contemporary imagery markedly higher for the
mono-dominant forest types compared to all others (Appendix Figure 4).
Provided that the young-regrowth and fallow permanent sampling plots
have seen recent disturbance the Wilcoxon signed rank test on the mixed
and mono-dominant plots between the historical and contemporary PC
values did not show a significant difference (p \textgreater{} 0.05).
Similarly, no significant different using PC values extracted from the
whole scene analysis was noted (p \textgreater{} 0.05). Any
relationships between contemporary Geo-Eye data and permanent sampling
plot measurements of Above Ground Carbon, stem density and species
richness were non-significant (p \textgreater{} 0.05, Appendix Figures
4-6).

Furthermore, visual inspection of the scene wide analysis suggests
historical scences do not show landscape wide canopy features (Figure 10
A-B), unlike the contemporary scene (Figure 10 C-D). In the contemporary
scene the FOTO algorithm picks up landscape features such as changes in
texture along the river valley (the diagonal line in Figure 10D).
However, no corresponding landscape patterns are found by the FOTO
algorithm in the historical orthomosaic.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/foto_maps}

}

\caption{RGB visualizations of the first three principal components of scene wide FOTO texture analysis of historical and current (Geo-Eye) imagery. Current permanent sampling plots of mono-dominant, mixed, fallow and young (edge) forest plots are marked with open triangles, open circles, open squares and crosses, respectively.}\label{fig:unnamed-chunk-14}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Finely grained spatial data sources, such as remote sensing imagery, are
rare before the satellite era (\textless{}1972). This lack of data
limits our understanding of how forest structure has varied over longer
time periods in remote areas. Long term assessment can be extended by
using large inventories of historical aerial survey data
\citep{kadmon1999, song2015, nita2018}. Despite the difficulties in
recovering this data and its limitations, such as invisible disturbances
\citep{peres2006}, remote sensing generally remains the best way to map
and quantify LULCC \citep{houghton2012}. In our study we used novel
numerical remote sensing techniques to valorize, for the first time,
historical remote sensing data in order to quantify (long term) land-use
and land-cover change and canopy structural properties in the central
Congo Basin. Despite these successes some methodological and research
considerations remain.

\hypertarget{methodological-considerations}{%
\subsection{Methodological
considerations}\label{methodological-considerations}}

\hypertarget{data-recovery-challenges}{%
\subsubsection{Data recovery
challenges}\label{data-recovery-challenges}}

In our study the archive data recovered was limited to contact prints
and therefore did not represent the true resolution of the original
negative. In addition, analogue photography clearly produces a distinct
softness compared to digital imagery (Figure 8). Despite favourable
nadir image acquistions \citep{verhoeven2017} image softness combined
with illumination effects between flight paths, and the self-similar
nature of vast canopy expanses \citep{park2019, simini2010, sole1995},
limited our ability to provide wall-to-wall coverage of the entire
dataset containing 334 images. Few man made features in the scenes also
made georeferencing challenging. Although the village of Yangambi
provided a range of buildings as (hard-edge) references other areas
within the central Congo Basin might have fewer permanent structures and
would require the use of soft-edged landscape features (e.g.~trees,
river outflows). Research has shown that soft-edged features can help
georeference scenes even when containing few man-made features
\citep{hughes2006}. Our two step georeferencing approach resulted in a
referencing accuracy of \textasciitilde{}4.7 \(\pm\) 4.3 m across
reference points. However, it shoud be noted that referencing accuracy
of the final scene is less constrained toward the edges of the scene.

\hypertarget{lulc-classification-validation}{%
\subsubsection{LULC classification \&
validation}\label{lulc-classification-validation}}

When classifying the orthomosaic into forest and non-forest states we
favoured a deep learning supervised classification using a CNN over
manual segmentation to guarantee an ``apples-to-apples'' comparison
between the historical and the contemporary GFC forest cover map used in
our analysis. We acknowledge that both the CNN and GFC land-use and
land-cover maps use different underlying features, i.e.~spatial or
spectral data, yet attain a similarly high accuracy of up to 99\%
\citep{hansen2013}. More so, when validating our CNN classifier against
GFC data (Figure 7) for a contemporary high resolution Geo-Eye
panchromatic image we reach an accuracy of \textasciitilde{}87\%,
despite a time difference of almost 60 years. Visual inspection of the
classification data in Figure 7 suggests that the GFC map more often
than not classifies non-forest areas as forest. Actual classification
accuracy of our algorithm might therefore be higher than our reported
value.

\hypertarget{scaling-opportunities}{%
\subsubsection{Scaling opportunities}\label{scaling-opportunities}}

Our approach uses broadly defined homogeneous polygons to construct a
balanced dataset of synthethic landscapes. The methodology is analoguous
to the use of sparse labelling as used by \citet{buscombe2018} and
contrasts with the standard methodologies which generally extract pixel
(windows) \citep{song2015} or delineate land cover classes
\citep{nita2018} to drive a classifier or analysis. More so, the use of
heavy image augmentation during model training sidesteps texture
representation issues which affect classification of image scenes with
inconsistent illumination or sharpness \citep{hudak1998} or ad-hoc
feature engineering \citep{song2015}. The use of synthetic landscapes
allowed us to account for most, but not all, of the variability within
our orthomosaic. Our analysis has shown that despite being trained on
historical data our model could map contemporary forest cover in remote
sensing data with similar spatial and spectral characteristics (Figure
7), suggesting that the classifier consistently works across both space
and time. We acknowledge that the use of synthetic landscapes is limited
by the available homogeneous areas to sample from and the number of
classes. Yet, the latter should not be a constrained as previous
research efforts have focussed on simple forest cover maps
\citep{hansen2013}.

\hypertarget{research-context}{%
\subsection{Research context}\label{research-context}}

\hypertarget{long-term-changes-in-lulc-and-above-ground-carbon-1}{%
\subsubsection{Long term changes in LULC and Above Ground
Carbon}\label{long-term-changes-in-lulc-and-above-ground-carbon-1}}

Our analysis shows that the majority of deforestation around Yangambi
happened toward the late 1950s (\textasciitilde{}16 200 ha).
Considerable reforestation has occured since the aerial survey was
executed (\textasciitilde{}9918 ha), and socio-economic instability
prevented further large scale forest exploitation. In particular, many
plantations have reached maturity and forest has re-established in
previously cleared or disturbed areas. The majority of this
reforestation takes the form of isolated patches of forest but is offset
by further deforestation of previously untouched forest. Generally, the
function and structure of forests can be influenced by forest edges that
are located up to 1 km away however most effects are pronounced within
the first 300 m from the edge \citep{gascon2000}. Our analysis of edge
effects on AGC has shown that the influence is negligible 200 m away
from the edge. \citet{phillips2006} have shown similar weak responses to
edge effects in the Amazon forest. Due to a lack of data on the extent
(depth) of edge effects and their influence on AGC beyond 200 m we did
not include any estimates of carbon loss or gain within these zones.
However, it must be stated that edges throughout the landscape make up a
substantial area and account for 13 151 ha. Thus edges could have a
substantial negative \citep{brinck2017} or positive \citep{reinmann2017}
influence on AGC. Similarly, uncertainties in how to explicitly correct
for plantations in the landscape present a further challenge. Similarly,
variability across mixed forest plots used in scaling above ground
carbon estimates due to deforestation introduced additional uncertainty
(see Appendix Table 3). Thus, although our estimates are only indicative
they do underscore the important influence of landscape structure in
carbon accounting. However, our findings do not indicate that
deforestation in Congo basin is declining, on the contrary.

Over the past half century there has been a clear shift in land use in
Yangambi (Figure 5). Land use has shifted away from a regular (fishbone)
deforestation pattern that emerges when (large scale) agricultural
interests dominate the landscape \citep{arima2015}, to a more fragmented
landscape (Figure 5D). The former is consistent with historical land
management at the time of the aerial survey \citep{bustillo2018}. These
regular patterns reversed due to a decrease in large scale intensive
agriculture and an increase in ad-hoc small scale subsistence farming
with large perimeter to area relationships (i.e.~ragged edges).
Consequently, edge effects in the current landscape are far more
pronounced than in the historical scene.

Visual inspection of the images also suggests that reforestation within
the historically cleared areas and experimental plots is not necessarily
limited to areas far removed from more densely populated areas. For
example, large reforested areas exist close to the Congo stream and
Yangambi village itself (Fig. 5). Here, regional political components,
such as land leases and large scale ownership could have played a role
in safeguarding some of these areas for rewilding or sustainable
management \citep{arima2014, larson2011}. Despite widespread
anthropogenic influences throughout the tropics \citep{lewis2015} the
retention of these forested areas show the potential of explicit or
implicit protective policy measures (e.g.~INERA concessions,
\citet{bustillo2018}) on a multi-decadal time scale. Reforestation in
non-continuous areas within Yangambi could increase landscape
connectivity and help increase biodiversity \citep{vandeperre2018}.

Our analysis therefore provides an opportunity to highlight and study
those regions that have previously suffered confirmed long-term
disturbances, and those that have been restored since. Assessments of
old plantations and recovering clear-cut forests can serve as a guide to
help estimate carbon storage capacity and forest recovery rates in
managed and unmanaged conditions
\citep{gourlet-fleury2013, sader1988, achard2014} over the mid- to
long-term, in support of rewilding and general forest restoration
\citep{arima2014, larson2011, vandeperre2018}. In addition, mapping
long-term edge effects can further support research into issues such as
receding forest edges \citep{gascon2000}.

\hypertarget{canopy-structure-foto-texture-analysis-2}{%
\subsubsection{Canopy structure \& FOTO texture
analysis}\label{canopy-structure-foto-texture-analysis-2}}

Finally, the FOTO technique used to quantify relationships between
canopy structure and forest characteristics rendered no valuable
insights of either the historical orthomosaic or recent Geo-Eye scene.
Similarly weak correlations were found previously by
\citet{solorzano2018}. In contrast, site based texture metric statistics
did show correspondence between historical and contemporary satellite
imagery. None of them were either consistent or significant. Although
visual interpretation shows distinctly different canopy structures
(Figure 5) the differences in how resolution is defined and issues
related to image quality prevented us from quantifying these further.
Unlike large scale studies by \citet{Ploton2012} we could not scale this
technique to historical data. The successful use of our CNN
classification model on a contemporary remote sensing data does suggest
that texture can be used to consistently capture canopy properties 60
years apart. Differences in PC between forest types (e.g.~mono-dominant
vs.~mixed, Figure 9) corroborate that texture can serve as a basis for
LULC classification. However, inflexibility on part of the FOTO
technique in dealing with non-standardized (historical) data, or scaling
these results to AGC values, limits its use case. We advise that future
valorisation efforts should preferentially work from foto negatives (if
available) to ensure optimal data quality in resolution, contrast and
sharpness.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Given the impact of tropical forest disturbances on atmospheric carbon
emissions, biodiversity and ecosystem productivity accurate long term
reporting of LULCC is an imperative. Our analysis of historical aerial
survey images (1958) of the Central Congo Basin provides a window into
the state of the forest at the start of the anthropocene. The use of a
CNN based LULC classifier, using synthetic landscapes based image
augmentation, provides a robust semi-supervised solution which scales
across space and time, even for image scenes with inconsistent
illumination or sharpness. Combined with contemporary remote sensing
data we have shown that historical aerial survey data can be used to
quantify long-term changes in LULC and AGC. We showed a shift from
previously highly structured industrial deforestation of large areas for
plantation purposes, to discrete smallholder clearing for farming,
increasing landscape fragmentation but also opportunties for substantial
regrowth. Efforts to quantify canopy texture features and their link to
AGC had limited to no success. Our analysis provides insights into the
rate at which deforestation and reforestation has taken place over a
multi-decadal scale in the central Congo basin. As such, it provides a
useful historical context while interpreting past and ongoing forest
research in the area.

\hypertarget{additional-information-and-declarations}{%
\section{Additional Information and
Declarations}\label{additional-information-and-declarations}}

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

Hufkens et al.~(2019). A curated dataset of aerial survey images over
the central Congo Basin, 1958. Zenodo:
\url{https://doi.org/10.5281/zenodo.3547767}. All data not included in
the latter repository can be found bundled with the analysis code as
listed below. Proprietary datasets (i.e.~Geo-Eye data) are not shared,
but purchase order numbers allow for acquisition of these datasets to
ensure reproducibility.

\hypertarget{code-availability}{%
\subsection{Code availability}\label{code-availability}}

All analysis code is available as an R / python \citep{rcoreteam2019}
projects (\url{https://khufkens.github.io/orthodrc} \&
\url{https://khufkens.github.io/orthodrc_cnn/}). The analysis relied
heavily on the `raster' \citep{hijmans2019}, `RStoolbox'
\citep{leutner2019}, and `landscapemetrics' \citep{hesselbarth2019}
packages, while post-processing and plotting was facilitated by the
`tidyverse' ecosystem \citep{wickham2017}, `ggthemes'
\citep{arnold2019}, `scales' \citep{wickham2018} and `cowplot'
\citep{wilke2019}. Additional plotting elements were formatted or
provided by `sf' \citep{pebesma2018} and `rnaturalearth'
\citep{south2017} packages, respectively. I'm grateful for the
contributions to the scientific community by the developers of these
packages.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
%
% % Only for the journal Methods and Protocols:
% % If you wish to submit a video article, please do so with any other supplementary material.
% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This research was supported through the Belgian Science Policy office
COBECORE project (BELSPO; grant BR/175/A3/COBECORE). K.H. acknowledges
funding from the European Union Marie Skłodowska-Curie Action (project
number 797668). K.H. grateful for the support of T.d.H. volunteering his
time in scanning the images.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{K.H. conceived and designed the study, analyzed the data, prepared
figures, tables and authored the final draft of the manuscript. T.d.H.
scanned all image data. E.K. and T.d.H. provided plot based AGC
estimates. T.D., K.J., E.K, H.B., P.S., F.V.,S.M., M.A., J.V.D.B., H.V
and L.W. reviewed the final manuscript.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest. The founding sponsors had
no role in the design of the study; in the collection, analyses, or
interpretation of data; in the writing of the manuscript, an in the
decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{bibliography.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}